### 2019-06-21 doc/fast_ssa/paper.tex
add documentation for this package

%\documentclass[manuscript,endfloat]{geophysics}
\documentclass[revised,endfloat]{geophysics}
%\documentclass[twocolumn,endfloat]{geophysics}

\usepackage{xy}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{url}
\usepackage{rotating}
\usepackage{lscape}
\usepackage{color}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{amssymb}      
\usepackage{algpseudocode}
\usepackage{booktabs}
 \usepackage{tikz}
 \usetikzlibrary{matrix}

\usepackage[ruled,linesnumbered]{algorithm2e}
\SetKwProg{Fn}{Function}{}{}

\usepackage{mathtools}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

\setlength\parindent{0pt}

\setlength\heavyrulewidth{0.25ex}

\DeclareGraphicsExtensions{.pdf}
\DeclareMathAlphabet\mathbfcal{OMS}{cmsy}{b}{n}
\DeclareMathOperator*{\argmin}{arg\,min}

\begin{document}

\title{Structured matrix}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\address{
\footnotemark[1]Department of Physics,\\
University of Alberta, \\
Edmonton, Alberta, Canada. \\}

\author{Wenlei Gao\footnotemark[1]}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABSTRACT                                                                   %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
The documentation for fast MSSA and relationship between multi-dimensional convolution and structured matrix, such as Circulant, Toeplitz and Hankel matrix.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\citep{korobeynikov2009} proposed the computation and memory efficient method for MSSA.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Method}
We start from the definition of discrete fourier transform.

\subsection{Discrete Fourier transform (DFT)}
For a vector ${\bf s}$ length of $N$, the discrete Fourier transform is  defined as
\begin{equation}
\tilde{s}_k  = \sum_{l=0}^{N-1} e^{-i 2 \pi k l / N} s_l \,,
\label{eq1}
\end{equation}
where the index $k = 0, 1, 2, \cdots, N-1$. Suppose the vector ${\bf s}$ represent a time series length of $N$ with time sampling interval is $\Delta t$. So the frequency sampling interval $\Delta f$ can be computed as 
\begin{equation}
\Delta f  = 1/ \Delta t / N = \frac{1}{N \Delta t}
\label{eq2}
\end{equation}
The Fourier transform for this time series in continueous case can be expressed as
\begin{equation}
\tilde{s}(f)  = \int e^{-i 2 \pi f t} s(t) dt 
\label{eq3}
\end{equation}
In discrete case and assume the frequency $f$ can be expressed as $f = k \Delta f$ and using the expression in equation \ref{eq2}
\begin{equation}
\begin{split}
\tilde{s}_{k \Delta f}  &= \sum_{l=0}^{N-1} e^{-i 2 \pi \, k \Delta f \,  l \Delta t} s_{l \Delta t}  \\
\tilde{s}_{k \Delta f}  &= \sum_{l=0}^{N-1} e^{-i 2 \pi \, k \frac{1}{N \Delta t} \,  l \Delta t} s_{l \Delta t} \\
\tilde{s}_{k}  &= \sum_{l=0}^{N-1} e^{-i 2 \pi \, k l / N} s_l \,.
\end{split}
\label{eq4}
\end{equation}
The the last equation is exactly same as the equation \ref{eq1}. Through this derivation, we can understand the physical meaning of the Fourier coefficients obtained via discrete Fourier transform. Note that the first coefficient corresponding to $0$ frequency and so as to the time series. The first element of ${\bf s}$ is the sample at $0$ time.

The above operation can be expressed into the matrix-vector form. To simplify the definition of Fourier transform matrix, we define a variable $z = e^{-i \pi  / N}$, so the Fourier transform matrix, which is a $N$-by-$N$ square matrix and the element $F_{kl} = z^{kl}$. It can be expanded as
\begin{equation}
{\bf F}_N = \begin{bmatrix}
1 & 1 & \cdots & 1 \\
1 & z & \cdots & z^{N-1} \\
1 & z^2 & \cdots & z^{2(N-1)} \\
\vdots & \vdots & \ddots & \vdots \\
1 & z^{N-1} & \cdots & z^{(N-1)(N-1)} \\
\end{bmatrix}
\label{eq5}
\end{equation}
As ${\bf F}_N$ is a orthogonal (but not orthonormal) matrix, we can be easily get the its inverse matrix
\begin{equation}
{\bf F}_N^{-1} = \frac{1}{N} {\bf F}_N^H = \frac{1}{N} {\bf F}_N^{\ast} \,
\label{eq6}
\end{equation}
where the super-script $^H$ represent complex conjugate transpose and $\ast$ indicate conjugate transpose. As the matrix ${\bf F}_N$ is a symmetrical matrix, that's why we can get equation \ref{eq6}.

\subsection{Circulant matrix}
An $9 \times 9$ circulant matrix is defined as
\begin{equation}
{\bf C} = 
\begin{bmatrix}
c_1 & c_9 & c_8 & c_7 & c_6 & c_5 &  c_4 & c_3 & c_2 \\ 
c_2 & c_1 & c_9 & c_8 & c_7 & c_6 &  c_5 & c_4 & c_3 \\ 
c_3 & c_2 & c_1 & c_9 & c_8 & c_7 &  c_6 & c_5 & c_4 \\ 
c_4 & c_3 & c_2 & c_1 & c_9 & c_8 &  c_7 & c_6 & c_5 \\ 
c_5 & c_4 & c_3 & c_2 & c_1 & c_9 &  c_8 & c_7 & c_6 \\ 
c_6 & c_5 & c_4 & c_3 & c_2 & c_1 &  c_9 & c_8 & c_7 \\ 
c_7 & c_6 & c_5 & c_4 & c_3 & c_2 &  c_1 & c_9 & c_8 \\ 
c_8 & c_7 & c_6 & c_5 & c_4 & c_3 &  c_2 & c_1 & c_9 \\ 
c_9 & c_8 & c_7 & c_6 & c_5 & c_4 &  c_3 & c_2 & c_1 \\ 
\end{bmatrix}
\end{equation} 
A circulant matrix is fully specified by its first column ${\bf c} = \begin{bmatrix}
c_1 & c_2 &  c_3 &  c_4 &  c_5 &  c_6 &  c_7 &  c_8 &  c_9  \end{bmatrix}^T$, we know that the product between circulant matrix and a vector ${\bf v}$ length of $N$ is equivalent to the circular convolution between vector ${bf c}$ and ${\bf v}$, and the circular convolution can be computed efficiently via Fast Fourier transform (FFT). So we have
\begin{equation}
{\bf Cv} = {\bf F}_N^{-1} \left(  {\bf F}_N {\bf c} \circ {\bf F}_N {\bf v} \right)  =  {\bf F}_N^{-1} \left(  diag({\bf F}_N {\bf c}) {\bf F}_N {\bf v} \right) \,,
\end{equation}
where the symbol $\circ$ and $diag$ represents Hardmard (element-wise) multiplication and building diagonal matrix from a vector.

\subsection{Toeplitz matrix}
Toeplitz matrix is associated with linear convolution, suppose we have two vectors ${\bf s} = \begin{bmatrix}  s_1 & s_2 & s_3 & s_4 & s_5  \end{bmatrix}^T$ and ${\bf f} =  \begin{bmatrix}  f_1 & f_2 & f_3 & f_4 & f_5  \end{bmatrix}^T$,  the linear convolution of these two vectors is vector ${\bf r}$ length of $9$. We can represent the convolution into matrix-vector form as
\begin{equation}
{\bf r} = {\bf  h} \ast {\bf f} = \begin{bmatrix}
r_1 \\
r_2 \\
r_3 \\
r_4 \\
r_5 \\
r_6 \\
r_7 \\
r_8 \\
r_9 \\
\end{bmatrix}
= \begin{bmatrix}
s_1    & \quad & \quad & \quad & \quad  \\ 
s_2    & s_1    & \quad & \quad & \quad  \\ 
s_3    & s_2    & s_1 & \quad & \quad \\ 
s_4    & s_3 & s_2 & s_1 & \quad  \\ 
s_5    & s_4 & s_3 & s_2 & s_1  \\ 
\quad & s_5 & s_4 & s_3 & s_2 \\ 
\quad & \quad & s_5 & s_4 & s_3  \\ 
\quad & \quad & \quad & s_5 & s_4 \\ 
\quad & \quad & \quad & \quad & s_5  \\ 
\end{bmatrix}
\begin{bmatrix}
f_1 \\
f_2 \\
f_3 \\
f_4 \\
f_5 \\
\end{bmatrix}
\end{equation}
The efficient linear convolution usually padding $0$ to vectors make their length equal to $9$, after padding zeros, we get new vector $\hat{\bf s} = \begin{bmatrix}  s_1 & s_2 & s_3 & s_4 & s_5 & 0 & 0 & 0 & 0 \end{bmatrix}$ and  $\hat{\bf f} = \begin{bmatrix}  f_1 & f_2 & f_3 & f_4 & f_5 & 0 & 0 & 0 & 0 \end{bmatrix} $, the circular convolution between $\hat{\bf s}$ and $\hat{\bf f}$ can be expressed as
\begin{equation}
\hat{\bf r} = \hat{\bf s} \circledast \hat{\bf f} = \begin{bmatrix}
s_1 & 0     & 0     & 0    & 0     & s_5 &  s_4 & s_3 & s_2 \\ 
s_2 & s_1 & 0     & 0    & 0     & 0     &  s_5 & s_4 & s_3 \\ 
s_3 & s_2 & s_1 & 0    & 0     & 0     &  0    & s_5 & s_4 \\ 
s_4 & s_3 & s_2 & s_1 & 0     & 0     &  0 & 0 & s_5 \\ 
s_5 & s_4 & s_3 & s_2 & s_1 & 0     &  0 & 0 & 0 \\ 
0    & s_5  & s_4 & s_3 & s_2 & s_1 &  0 & 0 & 0 \\ 
0    & 0     & s_5 & s_4  & s_3 & s_2 &  s_1 & 0 & 0 \\ 
0    & 0     & 0    & s_5  & s_4 & s_3 &  s_2 & s_1 & 0 \\ 
0    & 0     & 0    & 0     &  s_5 & s_4 &  s_3 & s_2 & s_1 \\ 
\end{bmatrix}
\begin{bmatrix}
f_1 \\
f_2 \\
f_3 \\
f_4 \\
f_5 \\
0 \\
0 \\
0 \\
0 \\
\end{bmatrix}
\label{eq10}
\end{equation}
When can see that $\hat{\bf r} = {\bf r}$, so linear convolution can be efficiently computed by padding zeros to vectors then compute circular convolution in frequency domain.

Let's consider another Toeplitz matrix built from a vector ${\bf c} = \begin{bmatrix}
c_1 & c_2 & c_3 & c_4 & c_5 & c_6 & c_7 & c_8 & c_9 \end{bmatrix}$ with length $N=9$, we first determine two integer: $L$ and $K$ make $L+K-1= N $. To make the matrix close to square form, we determine them by
\begin{equation}
L = \floor*{\frac{N}{2}}+1 ,\,\, K = N+1-L
\label{eq11}
\end{equation}
in this special case, $N=9$, so $L=K=5$, the Toeplitz matrix build from this vector can be expressed as
\begin{equation}
{\bf T} = 
\begin{bmatrix}
c_5 & c_4 & c_3 & c_2 & c_1 \\ 
c_6 & c_5 & c_4 & c_3 & c_2 \\ 
c_7 & c_6 & c_5 & c_4 & c_3 \\ 
c_8 & c_7 & c_6 & c_5 & c_4 \\ 
c_9 & c_8 & c_7 & c_6 & c_5 \\ 
\end{bmatrix}
\label{eq12}
\end{equation} 
The vector ${\bf c}$ can be uniquely determined from the last column and last row. There are two way to embedding this Teoplitz matrix into a circular matrix, the first to embedding this matrix at the lower-left part of the circular matrix created by vector ${\bf c}$, which shows as below as
\begin{center}
\begin{tikzpicture}
  \matrix (m)[
    matrix of math nodes,
    nodes in empty cells,
    minimum width=width("123"),
  ] {
c_1 & c_9 & c_8 & c_7 & c_6 & c_5 &  c_4 & c_3 & c_2 \\ 
c_2 & c_1 & c_9 & c_8 & c_7 & c_6 &  c_5 & c_4 & c_3 \\ 
c_3 & c_2 & c_1 & c_9 & c_8 & c_7 &  c_6 & c_5 & c_4 \\ 
c_4 & c_3 & c_2 & c_1 & c_9 & c_8 &  c_7 & c_6 & c_5 \\ 
c_5 & c_4 & c_3 & c_2 & c_1 & c_9 &  c_8 & c_7 & c_6 \\ 
c_6 & c_5 & c_4 & c_3 & c_2 & c_1 &  c_9 & c_8 & c_7 \\ 
c_7 & c_6 & c_5 & c_4 & c_3 & c_2 &  c_1 & c_9 & c_8 \\ 
c_8 & c_7 & c_6 & c_5 & c_4 & c_3 &  c_2 & c_1 & c_9 \\ 
c_9 & c_8 & c_7 & c_6 & c_5 & c_4 &  c_3 & c_2 & c_1 \\ 
  } ;
\draw (m-9-1.south west) rectangle (m-5-5.north east);
\end{tikzpicture}
\end{center}
The above circulant matrix is called ${\bf C}_1$. We can also create a new vector $\hat{\bf c} = \begin{bmatrix} c_5 & c_6 & c_7 & c_8 & c_9 & c_1 & c_2 & c_3 & c_4 \end{bmatrix}^T$, which is obtained by permutating the elements of the original vector, the Toeplitz matrix can be embedded at the upper-left part of the circulant matrix created from $\hat{\bf c}$ shows as bellow
 \begin{center}
\begin{tikzpicture}
  \matrix (m)[
    matrix of math nodes,
    nodes in empty cells,
    minimum width=width("123"),
  ] {
c_5 & c_4 & c_3 & c_2 & c_1 & c_9 &  c_8 & c_7 & c_6 \\ 
c_6 & c_5 & c_4 & c_3 & c_2 & c_1 &  c_9 & c_8 & c_7 \\ 
c_7 & c_6 & c_5 & c_4 & c_3 & c_2 &  c_1 & c_9 & c_8 \\ 
c_8 & c_7 & c_6 & c_5 & c_4 & c_3 &  c_2 & c_1 & c_9 \\ 
c_9 & c_8 & c_7 & c_6 & c_5 & c_4 &  c_3 & c_2 & c_1 \\ 
c_1 & c_9 & c_8 & c_7 & c_6 & c_5 &  c_4 & c_3 & c_2 \\ 
c_2 & c_1 & c_9 & c_8 & c_7 & c_6 &  c_5 & c_4 & c_3 \\ 
c_3 & c_2 & c_1 & c_9 & c_8 & c_7 &  c_6 & c_5 & c_4 \\ 
c_4 & c_3 & c_2 & c_1 & c_9 & c_8 &  c_7 & c_6 & c_5 \\ 
} ;
\draw (m-5-1.south west) rectangle (m-1-5.north east);
\end{tikzpicture}
\end{center}
 and call this new matrix as ${\bf C}_2$.

Suppose we need to compute the product between Toeplitz matrix ${\bf T}$ and a vector ${\bf v}$
\begin{equation}
{\bf r} = {\bf T}{\bf v} \,
\end{equation} 
where the length of ${\bf v}$ is $K$.
There are two way to do it efficiently via FFT, and these two ways corresponding to the two embedding methods. The first step is to padding vector ${\bf v}$ with $L-1$ $0s$ get a new vector $\hat{\bf v}$. Following above example, $K=L=5$ and ${\bf v} = \begin{bmatrix} v_1 & v_2 & v_3 & v_4 & v_5  \end{bmatrix}^T$, and the padded vector $\hat{\bf v} = \begin{bmatrix} v_1 & v_2 & v_3 & v_4 & v_5 & 0 & 0 & 0 & 0 \end{bmatrix}^T$, The product between ${\bf C}_1$ and $\hat{\bf v}$ are
\begin{equation}
\hat{\bf r}_1 = {\bf C}_1 \hat{\bf v} = \begin{bmatrix}
c_1 & c_9 & c_8 & c_7 & c_6 & c_5 &  c_4 & c_3 & c_2 \\ 
c_2 & c_1 & c_9 & c_8 & c_7 & c_6 &  c_5 & c_4 & c_3 \\ 
c_3 & c_2 & c_1 & c_9 & c_8 & c_7 &  c_6 & c_5 & c_4 \\ 
c_4 & c_3 & c_2 & c_1 & c_9 & c_8 &  c_7 & c_6 & c_5 \\ 
c_5 & c_4 & c_3 & c_2 & c_1 & c_9 &  c_8 & c_7 & c_6 \\ 
c_6 & c_5 & c_4 & c_3 & c_2 & c_1 &  c_9 & c_8 & c_7 \\ 
c_7 & c_6 & c_5 & c_4 & c_3 & c_2 &  c_1 & c_9 & c_8 \\ 
c_8 & c_7 & c_6 & c_5 & c_4 & c_3 &  c_2 & c_1 & c_9 \\ 
c_9 & c_8 & c_7 & c_6 & c_5 & c_4 &  c_3 & c_2 & c_1 \\ 
\end{bmatrix}
\begin{bmatrix}
v_1 \\
v_2 \\
v_3 \\
v_4 \\
v_5 \\
0 \\
0 \\
0 \\
0 \\
\end{bmatrix}
\end{equation}
where the last $L$ elements of $\hat{\bf r}_1$ equal to ${\bf r}$. On the other hand, we can take advantage of the circulate matrix ${\bf C}_2$, we have
\begin{equation}
\hat{\bf r}_2 = {\bf C}_2 \hat{\bf v} = \begin{bmatrix}
c_5 & c_4 & c_3 & c_2 & c_1 & c_9 &  c_8 & c_7 & c_6 \\ 
c_6 & c_5 & c_4 & c_3 & c_2 & c_1 &  c_9 & c_8 & c_7 \\ 
c_7 & c_6 & c_5 & c_4 & c_3 & c_2 &  c_1 & c_9 & c_8 \\ 
c_8 & c_7 & c_6 & c_5 & c_4 & c_3 &  c_2 & c_1 & c_9 \\ 
c_9 & c_8 & c_7 & c_6 & c_5 & c_4 &  c_3 & c_2 & c_1 \\ 
c_1 & c_9 & c_8 & c_7 & c_6 & c_5 &  c_4 & c_3 & c_2 \\ 
c_2 & c_1 & c_9 & c_8 & c_7 & c_6 &  c_5 & c_4 & c_3 \\ 
c_3 & c_2 & c_1 & c_9 & c_8 & c_7 &  c_6 & c_5 & c_4 \\ 
c_4 & c_3 & c_2 & c_1 & c_9 & c_8 &  c_7 & c_6 & c_5 \\ 
\end{bmatrix}
\begin{bmatrix}
v_1 \\
v_2 \\
v_3 \\
v_4 \\
v_5 \\
0 \\
0 \\
0 \\
0 \\
\end{bmatrix}
\end{equation}
So the first $L$ elements of $\hat{\bf r}_2$ equal to ${\bf r}$. We summarize the efficient computation of the product between Toeplitz matrix and a vector in Algorithm 1 and Algorithm 2
\begin{algorithm}
      \SetAlgoLined
      \Fn{ ${\bf v}= ttv({\bf c}, {\bf v}) $}{
      length of ${\bf c}$: $N = length({\bf c})$   \\
      compute $L = \floor*{\frac{N}{2}}$, $K = N+1 - L$   \\
      padding $L-1$ $0$s to the end of ${\bf v}$ to get a new vector $\hat{\bf v}$ with length of $N$\\
      ${\bf r}_1 = {\bf F}_N^{-1} \left(   {\bf F}_N {\bf c} \circ {\bf F}_N  \hat{\bf v}\right) $ \\
      take the last $L$ elements from ${\bf r}_1$ 
      }
      \caption{Teoplitz-Times-Vector}
\end{algorithm}

Another method is summarized in Algorithm 2 as 
\begin{algorithm}
      \SetAlgoLined
      \Fn{ ${\bf v}= ttv({\bf c}, {\bf v}) $}{
      length of ${\bf c}$: $N = length({\bf c})$   \\
      compute $L = \floor*{\frac{N}{2}}$, $K = N+1 - L$   \\
      padding $L-1$ $0$s to the end of ${\bf v}$ to get a new vector $\hat{\bf v}$ with length of $N$\\
      move the first $L-1$ element of ${\bf c}$ to the end, get a new vector $\hat{\bf c}$
      ${\bf r}_2 = {\bf F}_N^{-1} \left(   {\bf F}_N \hat{\bf c} \circ {\bf F}_N  \hat{\bf v}\right) $ \\
      take the first $L$ elements from ${\bf r}_2$ 
      }
      \caption{Teoplitz-Times-Vector}
\end{algorithm}

\bibliographystyle{seg}
\bibliography{paper}

\end{document}

### 2019-06-21 examples/arcis.jl

@@ -74,6 +74,8 @@ function stack_raw_data(path_fold::Ts, path_data::Ts) where {Ts<:String}
     # read the foldmap
     fold = read(path_fold)
 
+    
+
         dout = zeros(Float32, nt, numcr, numin)
     fdata = open(pd, "r")
     esize = 4

### 2019-06-21 src/mssa/fast_mssa.jl

@@ -1,0 +1,101 @@
+function build_circulant_matrix(v::Vector{Tv}) where {Tv<:Number}
+
+    # length of the vector
+    N = length(v)
+
+    # allocate space for circulant matrix N × N
+    C = Matrix{Tv}(undef, N, N)
+
+    # loop over column
+    for i2 = 1 : N
+
+        # loop over row, first part
+        for i1 = 1 : i2-1
+            idx = N - i2 + i1 + 1
+            C[i1,i2] = v[idx]
+        end
+
+        # loop over row, second part
+        for i1 = i2 : N
+            idx = i1 - i2 + 1
+            C[i1,i2] = v[idx]
+        end
+    end
+
+    return C
+end
+
+# test fast multiplication
+N = 111;
+s = randn(N);
+v = randn(N);
+C = build_circulant_matrix(s);
+
+r1 = C * v
+r2 = real(ifft( fft(s) .* fft(v) ))
+
+function build_toeplitz_matrix(v::Vector{Tv}) where {Tv<: Number}
+
+    # length of the vector
+    N = length(v)
+
+    # compute L, K
+    L = floor(Int64, N/2) + 1
+    K = N + 1 - L
+
+    # allocate space for Toeplitze matrix
+    T = Matrix{Tv}(undef, L, K)
+
+    # loop over column
+    for i2 = 1 : K
+
+        # loop over row
+        for i1 = 1 : L
+            idx = K + i1 - i2
+            T[i1,i2] = v[idx]
+        end
+    end
+
+    return T
+end
+
+# compute toeplitz times a vector
+function toeplitz_times_vector(c::Vector{Tv}, v::Vector{Tv}) where {Tv<:Number}
+
+    # length of vector
+    N = length(c)
+
+    # compute L, K
+    L = floor(Int64, N/2) + 1
+    K = N + 1 - L
+
+    # check the length of v
+    length(v) == K || DimensionMismatch()
+
+    # padding zeros to v
+    v_pad = vcat(v, zeros(Tv,L-1))
+
+    # fast computation
+    r_pad = real(ifft( fft(c) .* fft(v_pad) ) )
+
+    return r_pad[K:end]
+
+    # # the second way to compute
+    # c_hat = Vector{Tv}(undef, N)
+    # c_hat[L+1:N] .= c[1:K-1]
+    # c_hat[1:L]   .= c[K:N]
+    #
[…]

### 2019-06-28 doc/fast_ssa/paper.tex
working on the documentation

@@
 A circulant matrix is fully specified by its first column ${\bf c} = \begin{bmatrix}
-c_1 & c_2 &  c_3 &  c_4 &  c_5 &  c_6 &  c_7 &  c_8 &  c_9  \end{bmatrix}^T$, we know that the product between circulant matrix and a vector ${\bf v}$ length of $N$ is equivalent to the circular convolution between vector ${bf c}$ and ${\bf v}$, and the circular convolution can be computed efficiently via Fast Fourier transform (FFT).
+c_1 & c_2 &  c_3 &  c_4 &  c_5 &  c_6 &  c_7 &  c_8 &  c_9  \end{bmatrix}^T$, we know that the product between circulant matrix and a vector ${\bf v}$ length of $N$ is equivalent to the circular convolution between vector ${\bf c}$ and ${\bf v}$, and the circular convolution can be computed efficiently via Fast Fourier transform (FFT).
 So we have
 \begin{equation}
-{\bf Cv} = {\bf F}_N^{-1} \left(  {\bf F}_N {\bf c} \circ {\bf F}_N {\bf v} \right)  =  {\bf F}_N^{-1} \left(  diag({\bf F}_N {\bf c}) {\bf F}_N {\bf v} \right) \,,
+{\bf Cv} = {\bf F}_N^{-1} \left(  {\bf F}_N {\bf c} \circ {\bf F}_N {\bf v} \right)  =  {\bf F}_N^{-1} \left(  diag({\bf F}_N {\bf c}) \cdot {\bf F}_N {\bf v} \right) \,,
 \end{equation}
@@
 The above circulant matrix is called ${\bf C}_1$.
-We can also create a new vector $\hat{\bf c} = \begin{bmatrix} c_5 & c_6 & c_7 & c_8 & c_9 & c_1 & c_2 & c_3 & c_4 \end{bmatrix}^T$, which is obtained by permutating the elements of the original vector, the Toeplitz matrix can be embedded at the upper-left part of the circulant matrix created from $\hat{\bf c}$ shows as bellow
+We can create a new vector $\hat{\bf c} = \begin{bmatrix} c_5 & c_6 & c_7 & c_8 & c_9 & c_1 & c_2 & c_3 & c_4 \end{bmatrix}^T$, which is obtained by permutating the elements of the original vector, the Toeplitz matrix can be embedded at the upper-left part of the circulant matrix created from $\hat{\bf c}$ shows as bellow
  \begin{center}
@@
 There are two way to do it efficiently via FFT, and these two ways corresponding to the two embedding methods.
-The first step is to padding vector ${\bf v}$ with $L-1$ $0s$ get a new vector $\hat{\bf v}$.
+The first step is to padding vector ${\bf v}$ with $L-1$ $0$s get a new vector $\hat{\bf v}$.
 Following above example, $K=L=5$ and ${\bf v} = \begin{bmatrix} v_1 & v_2 & v_3 & v_4 & v_5  \end{bmatrix}^T$, and the padded vector $\hat{\bf v} = \begin{bmatrix} v_1 & v_2 & v_3 & v_4 & v_5 & 0 & 0 & 0 & 0 \end{bmatrix}^T$, The product between ${\bf C}_1$ and $\hat{\bf v}$ are
@@
       length of ${\bf c}$: $N = length({\bf c})$   \\
-      compute $L = \floor*{\frac{N}{2}}$, $K = N+1 - L$   \\
+      compute $L = \floor*{\frac{N}{2}}+1$, $K = N+1 - L$   \\
       padding $L-1$ $0$s to the end of ${\bf v}$ to get a new vector $\hat{\bf v}$ with length of $N$\\
@@
       length of ${\bf c}$: $N = length({\bf c})$   \\
-      compute $L = \floor*{\frac{N}{2}}$, $K = N+1 - L$   \\
+      compute $L = \floor*{\frac{N}{2}}+1$, $K = N+1 - L$   \\
       padding $L-1$ $0$s to the end of ${\bf v}$ to get a new vector $\hat{\bf v}$ with length of $N$\\
-      move the first $L-1$ element of ${\bf c}$ to the end, get a new vector $\hat{\bf c}$
+      move the first $K-1$ element of ${\bf c}$ to the end, get a new vector $\hat{\bf c}$
       ${\bf r}_2 = {\bf F}_N^{-1} \left(   {\bf F}_N \hat{\bf c} \circ {\bf F}_N  \hat{\bf v}\right) $ \\
@@
 
+\subsection{Hankel matrix}
+Hankel matrix of size $L \times K$ created from the same vector ${\bf c}$ is given as
+\begin{equation}
+{\bf H} = 
+\begin{bmatrix}
+c_1 & c_2 & c_3 & c_4 & c_5 \\ 
+c_2 & c_3 & c_4 & c_5 & c_6 \\ 
+c_3 & c_4 & c_5 & c_6 & c_7 \\ 
+c_4 & c_5 & c_6 & c_7 & c_8 \\ 
+c_5 & c_6 & c_7 & c_8 & c_9 \\ 
+\end{bmatrix}
+\label{eq16}
+\end{equation}
+which can be determined from the first column and last row.
+
+\textcolor{red}{Note: The circulant matrix and Teoplitz matrix are constant on the diagonals. However, Hankel matrix are constant on the anti-diagonals.}
+
+One can convert a Hankel matrix to a Toeplitz matrix by reversing its columns.
+The columns reversing process can be represented by right multiplying a $K \times k$ matrix ${\bf R}$, whose elements are $1$ on the main anti-diagonals and 0 elsewhere.
+Follow the toy example, we can get
+\begin{equation}
+{\bf T}= \begin{bmatrix}
+c_5 & c_4 & c_3 & c_2 & c_1 \\ 
+c_6 & c_5 & c_4 & c_3 & c_2 \\ 
+c_7 & c_6 & c_5 & c_4 & c_3 \\ 
+c_8 & c_7 & c_6 & c_5 & c_4 \\ 
+c_9 & c_8 & c_7 & c_6 & c_5 \\ 
+\end{bmatrix} = {\bf HR} = 
+\begin{bmatrix}
+c_1 & c_2 & c_3 & c_4 & c_5 \\ 
+c_2 & c_3 & c_4 & c_5 & c_6 \\ 
+c_3 & c_4 & c_5 & c_6 & c_7 \\ 
+c_4 & c_5 & c_6 & c_7 & c_8 \\ 
+c_5 & c_6 & c_7 & c_8 & c_9 \\ 
+\end{bmatrix}
+\begin{bmatrix}
+\quad & \quad &  \quad &  \quad & 1 \\
+\quad & \quad &  \quad &  1 & \quad \\
+\quad & \quad &  1 &  \quad & \quad \\
+\quad & 1 &  \quad &  \quad & \quad \\
+1 & \quad &  \quad &  \quad & \quad \\
+\end{bmatrix}
+\end{equation}
+
+Before we discuss about the multiplication between Hankel matrix and a vector, we first look at the property of the column reversing matrix ${\bf R}$, we have 
+\begin{equation}
+{\bf R} = {\bf R}^T = {\bf R}^{-1}
+\end{equation}
+So the product between Hankel matrix ${\bf H}$ and a vector ${\bf v}$ is given as 
+\begin{equation}
+{\bf r} = {\bf Hv} = {\bf HRR}^{-1}{\bf v} = {\bf HRRv} = ({\bf HR})({\bf Rv}) = {\bf T} \hat {\bf v}
+\end{equation}
+where $\bf{v}$ is the reversed version of the vector ${\bf v}$.
+
 \bibliographystyle{seg}

### 2019-06-28 src/mssa/fast_mssa.jl

@@ -1,3 +1,6 @@
+"""
+   build a circulant matrix from a vector
+"""
 function build_circulant_matrix(v::Vector{Tv}) where {Tv<:Number}
 
     # length of the vector
@@ -25,16 +28,11 @@ function build_circulant_matrix(v::Vector{Tv}) where {Tv<:Number}
     return C
 end
 
-# test fast multiplication
-N = 111;
-s = randn(N);
-v = randn(N);
-C = build_circulant_matrix(s);
-
-r1 = C * v
-r2 = real(ifft( fft(s) .* fft(v) ))
-
-function build_toeplitz_matrix(v::Vector{Tv}) where {Tv<: Number}
+"""
+   build a Toeplitze matrix from a vector, this matrix is a square (length of v is odd)
+or one more row than columns (length of v is even)
+"""
+function build_toeplitz_matrix(v::Vector{Tv}) where {Tv <: Number}
 
     # length of the vector
     N = length(v)
@@ -59,8 +57,37 @@ function build_toeplitz_matrix(v::Vector{Tv}) where {Tv<: Number}
     return T
 end
 
+"""
+   build a Hankel matrix from a vector, this matrix is a square (length of v is odd)
+or one more row than columns (length of v is even)
+"""
+function build_hankel_matrix(v::Vector{Tv}) where {Tv <: Number}
+
+    # length of the vector
+    N = length(v)
+
+    # compute L, K
+    L = floor(Int64, N/2) + 1
+    K = N + 1 - L
+
+    # allocate space for Toeplitze matrix
+    H = Matrix{Tv}(undef, L, K)
+
+    # loop over column
+    for i2 = 1 : K
+
+        # loop over row
+        for i1 = 1 : L
+            idx = i1+i2-1
+            H[i1,i2] = v[idx]
+        end
+    end
+
+    return H
+end
+
 # compute toeplitz times a vector
-function toeplitz_times_vector(c::Vector{Tv}, v::Vector{Tv}) where {Tv<:Number}
+function toeplitz_times_vector(c::Vector{Tv}, v::Vector{Tv}; embed_flag=1) where {Tv<:Number}
 
     # length of vector
     N = length(c)
@@ -76,26 +103,67 @@ function toeplitz_times_vector(c::Vector{Tv}, v::Vector{Tv}) where {Tv<:Number}
[…]

### 2019-07-04 doc/fast_ssa/paper.tex
review the property of Fourier transform

@@
  \usetikzlibrary{matrix}
+ \DeclareMathAlphabet\mathbfcal{OMS}{cmsy}{b}{n}
 
@@
 where the super-script $^H$ represent complex conjugate transpose and $\ast$ indicate conjugate transpose. As the matrix ${\bf F}_N$ is a symmetrical matrix, that's why we can get equation \ref{eq6}.
+
+\subsection{Properties of Fourier transform}
+
+In this section, we review the properties of discrete Fourier transform ({\bf DFT}), most of the properties are derived in continuous domain as it is much easier for the derivation, but the discrete version are verified by $Julia$ code.
+
+We define the Fourier transform as
+\begin{equation}
+X(\omega) = \int_{-\infty}^{\infty} x(t) e^{-i\omega t} dt
+\end{equation}
+where $\omega$ is radian frequency and $t$ is time. The above operation can be simplified as 
+\begin{equation}
+{\mathbfcal F}[x(t)] \rightarrow X(\omega)
+\end{equation}
+
+\subsubsection{1. time shift}
+\begin{equation}
+{\mathbfcal F}[x(t-t_0)] = X(\omega)e^{i \omega t_0}
+\end{equation}
+
+{\bf Proof}:
+$$
+{\mathbfcal F}[x(t-t_0)]  = \int_{-\infty}^{\infty} x(t - t_0) e^{-i\omega t} dt  
+$$
+we set $t^{\prime} = t - t_0$ and exchange the variables, we can get
+$$
+{\mathbfcal F}[x(t-t_0)]  = \int_{-\infty}^{\infty} x(t^{\prime}) e^{-i\omega (t^{\prime} + t_0)} d t^{\prime}  $$
+$$
+{\mathbfcal F}[x(t-t_0)] = e^{i \omega t_0} \int_{-\infty}^{\infty} x(t^{\prime}) e^{-i\omega t^{\prime}} d t^{\prime} = X(\omega) e^{i \omega t_0}
+$$
+
+\subsubsection{2. frequency modulation}
+\begin{equation}
+{\mathbfcal F}[x(t)e^{i \omega_0 t}] = X(\omega-\omega_0)
+\end{equation}
+{\bf Proof}:
+$$
+{\mathbfcal F}[x(t)e^{i \omega_0 t}]  = \int_{-\infty}^{\infty} x(t)e^{i \omega_0 t} e^{-i\omega t} dt  \\
+=  \int_{-\infty}^{\infty} x(t) e^{-i (\omega-\omega_0) t} dt  =  X(\omega-\omega_0)
+$$
+\textcolor{red}{Note: this is the property we forget very often.}
+
+\subsubsection{3. time reversal}
+\begin{equation}
+{\mathbfcal F}[x(-t)] = X(-\omega)
+\end{equation}
+{\bf Proof}:
+$$ {\mathbfcal F}[x(-t)]  = \int_{-\infty}^{\infty} x(-t)e^{-i \omega t} dt  $$
+Set $t^{\prime} = -t$ and change the variables we can get
+$$ {\mathbfcal F}[x(-t)]  = \int_{\infty}^{-\infty} x(t^{\prime}) e^{-i \omega (-t^{\prime})} d (-t^{\prime}) \\
+= \int_{-\infty}^{\infty} x(t^{\prime}) e^{-i (-\omega) t^{\prime}} d t^{\prime} \\
+= X(-\omega)$$
+Based on this property, we can get two extensions, if the time function is and even function, that is
+$$x(t) = x(-t) \rightarrow X(\omega) = X(-\omega)$$
+if the time function is odd function, the Fourier transform of this function is also an odd function
+$$x(t) = -x(-t) \rightarrow X(\omega) = - X(-\omega)$$
+
+\subsubsection{4. time and frequency stretching}
+\begin{equation}
+{\mathbfcal F}[x(at)] = \frac{1}{a}X(\frac{\omega}{a})
+\end{equation}
+where $a$ is positive number
+
+{\bf Proof}:
+$$ {\mathbfcal F}[x(at)]  = \int_{-\infty}^{\infty} x(at)e^{-i \omega t} dt  $$
+Set $t^{\prime} = a t$, By change the variable, we can get
+$$ {\mathbfcal F}[x(at)]  = \int_{-\infty}^{\infty} x(t^{\prime})e^{-i \omega \frac{t^{\prime}}{a}} d\frac{t^{\prime}}{a} = \frac{1}{a}  \int_{-\infty}^{\infty} x(t^{\prime}) e^{-i \frac{\omega}{a} t^{\prime}} d t^{\prime} = \frac{1}{a}X{(\frac{w}{a})} $$
+\textcolor{red}{Stretching in one domain means squeezing in another domain and vis versa.}
+
+\subsubsection{5. Complex conjugate}
+\begin{equation}
+{\mathbfcal F}[x^*(t)] = X^*(- \omega)
+\end{equation}
+where the super-script $^*$ represent complex conjugate.
+
+{\bf Proof}:
+$$\int_{-\infty}^{\infty}  x^*(t) e^{-i \omega t} dt = \int_{-\infty}^{\infty} \left( x(t) e^{i \omega t}\right)^* dt = \int_{-\infty}^{\infty} \left(x(t) e^{-i (-\omega) t}\right)^* dt = \left( \int_{-\infty}^{\infty} x(t) e^{-i (-\omega) t} dt\right)^* = X^*(-\omega)  $$
 
@@
       \SetAlgoLined
-      \Fn{ ${\bf v}= ttv({\bf c}, {\bf v}) $}{
+      \Fn{ ${\bf r}= ttv({\bf c}, {\bf v}) $}{
       length of ${\bf c}$: $N = length({\bf c})$   \\
@@
 One can convert a Hankel matrix to a Toeplitz matrix by reversing its columns.
-The columns reversing process can be represented by right multiplying a $K \times k$ matrix ${\bf R}$, whose elements are $1$ on the main anti-diagonals and 0 elsewhere.
+The columns reversing process can be represented by right multiplying a $K \times K$ matrix ${\bf R}$, whose elements are $1$ on the main anti-diagonals and 0 elsewhere.
 Follow the toy example, we can get
@@
 where $\bf{v}$ is the reversed version of the vector ${\bf v}$.
+From above derivations, we can see that the multiplication between a Hankel matrix and a vector is converted to the multiplication between Teoplitz matrix and a vector.
+So the algorithm for the multiplication between Hankel matrix and a vector is summarized in Algorithm 3 as
+\begin{algorithm}
+      \SetAlgoLined
+      \Fn{ ${\bf r}= htv({\bf c}, {\bf v}) $}{
+      length of ${\bf c}$: $N = length({\bf c})$   \\
+      compute $L = \floor*{\frac{N}{2}}+1$, $K = N+1 - L$   \\
+      reverse the order of the elements of ${\bf v}$: $\hat{\bf v} = reverse({\bf v})$ \\
+      call the function Toeplitz-times-vector: ${\bf r} = ttv({\bf c}, \hat{\bf v})$ \\
+      \Return {\bf r}}
+      \caption{Hankel-Times-Vector}
+\end{algorithm}
 

### 2019-07-04 src/mssa/fast_mssa.jl

@@ -142,7 +142,7 @@ function hankel_times_vector(c::Vector{Tv}, v::Vector{Tv}; embed_flag=1) where {
     # reverse the order of the elements of v
     v_hat = Vector{Tv}(undef, K)
     for i = 1 : K
-        v_hat[i] = v[N-i+1]
+        v_hat[i] = v[K-i+1]
     end
 
     return toeplitz_times_vector(c, v_hat; embed_flag=embed_flag)
@@ -167,3 +167,15 @@ end
 # T  = build_toeplitz_matrix(c);
 # r1 = T * v;
 # norm(r-r1) / norm(r)
+
+
+
+# test Hankel matrix times a vector
+L = 51; K = 51;
+N = L+K-1;
+c = randn(N); v = randn(K);
+r = hankel_times_vector(c, v; embed_flag=2);
+
+H  = build_hankel_matrix(c);
+r1 = H * v;
+norm(r-r1) / norm(r)

### 2019-07-04 src/mssa/clean_old_work.jl
old work need to be up-dated

@@ -1,0 +1,1577 @@
+# ==============================================================================
+#                        real data example
+# ==============================================================================
+using Registration
+
+i3 = 24; i4 = 5; dt =0.004;
+pin = "/Users/wenlei/Desktop/ArcisData/SmallCube/PP_wave/patches/patch"
+pathin  = join([pin "_" "$i3" "_" "$i4" ".bin"]); fid = open(pathin, "r")
+nt  = read(fid, Int32); ntrace = read(fid, Int32);
+mark= read(fid, Int32, ntrace); mark = convert(Vector{Int64}, mark);
+dim = read(fid, Int32, 4); dim = convert(Vector{Int64}, dim);
+dobs= read(fid, Float32, nt*ntrace); dobs=reshape(dobs,Int64(nt),Int64(ntrace));
+close(fid)
+dori = zeros(Float32, nt, dim[1], dim[2], dim[3], dim[4]);
+for i = 1 : length(mark)
+    idx = ind2sub((dim[1],dim[2],dim[3],dim[4]), mark[i])
+    dori[:,idx[1], idx[2], idx[3], idx[4]] = dobs[:,i]
+end
+dori = reshape(dori, Int64(nt), dim[1], dim[2], dim[3], dim[4]);
+dori = permutedims(dori, (1,3,2,4,5));
+
+itl=376; nt=165; itu=itl+nt-1;
+dobs = convert(Array{Float64,5}, dori)[itl:itu, :,:,:,:];
+(nt, N1, N2, N3, N4) = size(dobs);
+(Amp, f) = freqSpectrum(reshape(dobs, nt, N1*N2*N3*N4), dt)
+
+Fd = fft(dobs, 1)
+f0 = 32.0; iw=floor(Int64, f0*nt*dt)+1;
+
+Indicator = ones(N1, N2, N3, N4)
+for i4 = 1 : N4
+    for i3 = 1 : N3
+        for i2 = 1 : N2
+            for i1 = 1 : N1
+                if norm(dobs[:,i1,i2,i3,i4]) == 0.0
+                   Indicator[i1,i2,i3,i4] = 0.0
+                end
+            end
+        end
+    end
+end
+
+# find the row and column index of missing trace embeded in hankel matrix
+L1 = floor(Int64, N1/2) + 1; K1 = N1 - L1 + 1;
+L2 = floor(Int64, N2/2) + 1; K2 = N2 - L2 + 1;
[…]

### 2019-07-05 doc/fast_ssa/paper.tex
try to finish ssa

@@
 \subsection{Hankel matrix}
-Hankel matrix of size $L \times K$ created from the same vector ${\bf c}$ is given as
+We divide Hankel matrices into two classes, one is the number of elements is even and the other one is odd.
+We first discuss about when the number of elements is even, for example $N=6$.
+The vector can be expressed as
+$${\bf c} = \begin{bmatrix}
+c_1\\
+c_2\\
+c_3\\
+c_4\\
+c_5\\
+c_6\\
+\end{bmatrix}$$ 
+We follow the $L = \floor*{\frac{N}{2}}+1 = 4$ and $K = N+1-L = 3$ to determine the number of rows and columns of the result Hankel matrix as
 \begin{equation}
@@
 \begin{bmatrix}
-c_1 & c_2 & c_3 & c_4 & c_5 \\ 
-c_2 & c_3 & c_4 & c_5 & c_6 \\ 
-c_3 & c_4 & c_5 & c_6 & c_7 \\ 
-c_4 & c_5 & c_6 & c_7 & c_8 \\ 
-c_5 & c_6 & c_7 & c_8 & c_9 \\ 
+c_1 & c_2 & c_3 \\ 
+c_2 & c_3 & c_4 \\ 
+c_3 & c_4 & c_5 \\ 
+c_4 & c_5 & c_6 \\ 
 \end{bmatrix}
-\label{eq16}
+\label{eq23}
 \end{equation}
-which can be determined from the first column and last row.
+We also give the Hermitian transpose (complex-conjugate transpose) of the Hankel matrix as
+\begin{equation}
+{\bf H}^H = 
+\begin{bmatrix}
+c_1^*  & c_2^*  & c_3^*  & c_4^*  \\ 
+c_2^*  & c_3^*  & c_4^*  & c_5^*  \\ 
+c_3^*  & c_4^*  & c_5^*  & c_6^*  \\ 
+\end{bmatrix}
+\label{eq24}
+\end{equation}
+where the super-script indicate Hermitian transpose.
+We can see that the Hermitian transpose of Hankel matrix is still an Hankel matrix. The vector elements can be determined from the first column and last row.
 
@@
 The columns reversing process can be represented by right multiplying a $K \times K$ matrix ${\bf R}$, whose elements are $1$ on the main anti-diagonals and 0 elsewhere.
-Follow the toy example, we can get
+Which given as
 \begin{equation}
 {\bf T}= \begin{bmatrix}
-c_5 & c_4 & c_3 & c_2 & c_1 \\ 
-c_6 & c_5 & c_4 & c_3 & c_2 \\ 
-c_7 & c_6 & c_5 & c_4 & c_3 \\ 
-c_8 & c_7 & c_6 & c_5 & c_4 \\ 
-c_9 & c_8 & c_7 & c_6 & c_5 \\ 
+c_3 & c_2 & c_1 \\ 
+c_4 & c_3 & c_2 \\ 
+c_5 & c_4 & c_3 \\ 
+c_6 & c_5 & c_4 \\ 
 \end{bmatrix} = {\bf HR} = 
 \begin{bmatrix}
-c_1 & c_2 & c_3 & c_4 & c_5 \\ 
-c_2 & c_3 & c_4 & c_5 & c_6 \\ 
-c_3 & c_4 & c_5 & c_6 & c_7 \\ 
-c_4 & c_5 & c_6 & c_7 & c_8 \\ 
-c_5 & c_6 & c_7 & c_8 & c_9 \\ 
+c_1 & c_2 & c_3  \\ 
+c_2 & c_3 & c_4  \\ 
+c_3 & c_4 & c_5  \\ 
+c_4 & c_5 & c_6  \\ 
 \end{bmatrix}
 \begin{bmatrix}
-\quad & \quad &  \quad &  \quad & 1 \\
-\quad & \quad &  \quad &  1 & \quad \\
-\quad & \quad &  1 &  \quad & \quad \\
-\quad & 1 &  \quad &  \quad & \quad \\
-1 & \quad &  \quad &  \quad & \quad \\
+\quad &  \quad & 1 \\
+\quad &  1 & \quad \\
+1 &  \quad & \quad \\
+\end{bmatrix}
+\end{equation}
+The columns reverse of the Hermitian transpose of Hankel matrix is also a Toeplitz matrix given as 
+\begin{equation}
+\hat{\bf T}= \begin{bmatrix}
+c_4^* & c_3^* & c_2^* & c_1^* \\ 
+c_5^* & c_4^* & c_3^* & c_2^* \\ 
+c_6^* & c_5^* & c_4^* & c_3^* \\ 
+\end{bmatrix} = {\bf H}^H{\bf R} = 
+\begin{bmatrix}
+c_1^*  & c_2^*  & c_3^*  & c_4^*  \\ 
+c_2^*  & c_3^*  & c_4^*  & c_5^*  \\ 
+c_3^*  & c_4^*  & c_5^*  & c_6^*  \\ 
+\end{bmatrix}
+\begin{bmatrix}
+\quad & \quad &  \quad & 1 \\
+\quad & \quad &  1 & \quad \\
+\quad & 1        &  \quad & \quad \\
+1        & \quad &  \quad & \quad \\
 \end{bmatrix}
@@
 
-Before we discuss about the multiplication between Hankel matrix and a vector, we first look at the property of the column reversing matrix ${\bf R}$, we have 
+Before we discuss about the multiplication between Hankel matrix or the Hermitian transpose and a vector, we first look at the property of the column reversing matrix ${\bf R}$, we have 
 \begin{equation}
@@
 \end{equation}
-where $\bf{v}$ is the reversed version of the vector ${\bf v}$.
-From above derivations, we can see that the multiplication between a Hankel matrix and a vector is converted to the multiplication between Teoplitz matrix and a vector.
-So the algorithm for the multiplication between Hankel matrix and a vector is summarized in Algorithm 3 as
+where $\bf{v}$ is the reversed version of the vector ${\bf v}$. Accordingly, the Hermitian transpose of Hankel matrix times a vector can be represented as
+\begin{equation}
+\hat{\bf r} = {\bf H}^H{v} = {\bf H}^H {\bf RR}^{-1}{\bf v} = {\bf H}^H{\bf RRv} = ({\bf H}^H{\bf R})({\bf Rv}) = \hat{\bf T} \hat {\bf v}
+\end{equation}
+There are two ways to embed ${\bf HR}$ and ${\bf H}^H {\bf R}$ into a circulant matrix, which is given as
+\begin{center}
+\begin{tikzpicture}
+  \matrix (m)[
+    matrix of math nodes,
+    nodes in empty cells,
+    minimum width=width("123"),
+  ] {
+c_1 & c_6 & c_5 & c_4 & c_3 & c_2 \\ 
+c_2 & c_1 & c_6 & c_5 & c_4 & c_3 \\ 
+c_3 & c_2 & c_1 & c_6 & c_5 & c_4 \\ 
+c_4 & c_3 & c_2 & c_1 & c_6 & c_5 \\ 
+c_5 & c_4 & c_3 & c_2 & c_1 & c_6 \\ 
+c_6 & c_5 & c_4 & c_3 & c_2 & c_1 \\ 
+} ;
+\draw (m-6-1.south west) rectangle (m-3-3.north east);
+\draw (m-6-1.south west) rectangle (m-4-4.north east);
+\end{tikzpicture}
+\end{center}
+The second way to embed a Hankel matrix into a circulant matrix is given as
+\begin{center}
+\begin{tikzpicture}
+  \matrix (m)[
+    matrix of math nodes,
+    nodes in empty cells,
+    minimum width=width("123"),
+  ] {
+c_3 & c_2 & c_1 & c_6 & c_5 & c_4 \\ 
+c_4 & c_3 & c_2 & c_1 & c_6 & c_5 \\ 
+c_5 & c_4 & c_3 & c_2 & c_1 & c_6 \\ 
+c_6 & c_5 & c_4 & c_3 & c_2 & c_1 \\ 
+c_1 & c_6 & c_5 & c_4 & c_3 & c_2 \\ 
+c_2 & c_1 & c_6 & c_5 & c_4 & c_3 \\ 
+} ;
+\draw (m-4-1.south west) rectangle (m-1-3.north east);
+\draw (m-6-3.south west) rectangle (m-4-6.north east);
+\end{tikzpicture}
+\end{center}
+
+Based on the two different embedding methods, we also have two algorithm to compute the product between Hankel matrix or Hermitian transpose and a vector,  which is given as
 \begin{algorithm}
       \SetAlgoLined
-      \Fn{ ${\bf r}= htv({\bf c}, {\bf v}) $}{
+      \Fn{ ${\bf r}= htv({\bf c}, {\bf v}; flag="forward" \,\,or\,\,"adjoint") $}{
       length of ${\bf c}$: $N = length({\bf c})$   \\
@@
       reverse the order of the elements of ${\bf v}$: $\hat{\bf v} = reverse({\bf v})$ \\
-      call the function Toeplitz-times-vector: ${\bf r} = ttv({\bf c}, \hat{\bf v})$ \\
-      \Return {\bf r}}
+      Fourier transform: $\hat{\bf c} = fft({\bf c})$ \\
+      \If{ Forward}{
+      Padding zeros to $\hat{\bf v}$:  $\hat{\bf v} = [\hat{\bf v}, zeros(L-1)]$     \\
+      Fourier transform: $\tilde{\bf v}  = fft(\hat{\bf v})$ \\
+      element-wise multiplication: ${\bf r} = ifft(\hat{\bf c} \circ \tilde{\bf v})$ \\
+      \Return ${\bf r}[K:N]$}
+      \If{ Adjoint }{
+      Padding zeros to $\hat{\bf v}$:  $\hat{\bf v} = [\hat{\bf v}, zeros(K-1)]$     \\
+      Fourier transform: $\tilde{\bf v}  = fft(\hat{\bf v})$ \\
+      Conjugate property of Fourier transform: $\tilde{\bf c}[1] = \hat{\bf c}^*[1]$ and  $\tilde{\bf c}[i] = \hat{\bf c}^*[N-i+2] $ \\
+      element-wise multiplication: ${\bf r} = ifft(\tilde{\bf c} \circ \tilde{\bf v})$ \\
+      \Return ${\bf r}[L:N]$}}
       \caption{Hankel-Times-Vector}
 \end{algorithm}
+\textcolor{red}{The expression in Line 15 use the conjugate property of Fourier transform, which is stated as}
+$$
+\mathbfcal{F}[x(t)] \rightarrow X(\omega)$$
+$$
+\mathbfcal{F}[x^*(t)] \rightarrow X^*(-\omega)
+$$
 

### 2019-07-05 src/mssa/fast_mssa.jl

@@ -58,33 +58,372 @@ function build_toeplitz_matrix(v::Vector{Tv}) where {Tv <: Number}
 end
 
 """
-   build a Hankel matrix from a vector, this matrix is a square (length of v is odd)
-or one more row than columns (length of v is even)
+   build a multi-level block Hankel matrix from a multi-dimensional array
 """
-function build_hankel_matrix(v::Vector{Tv}) where {Tv <: Number}
+function build_hankel_matrix(d::Array{Tv}) where {Tv <: Number}
 
-    # length of the vector
-    N = length(v)
+    # determine the dimensions of an array
+    dims = size(d)
+    N    = length(dims)
 
-    # compute L, K
-    L = floor(Int64, N/2) + 1
-    K = N + 1 - L
+    # level 1
+    if N == 1
+       L1 = floor(Int64,dims[1]/2)+1; K1 = dims[1]-L1+1
 
-    # allocate space for Toeplitze matrix
-    H = Matrix{Tv}(undef, L, K)
+       # allocate memory for Hankel matrix
+       H  = Array{Tv,2}(undef, L1, K1)
 
-    # loop over column
-    for i2 = 1 : K
+       # build hankel
+       for j1 = 1 : K1
+           for i1 = 1 : L1
+               n1 = i1+j1-1
+               H[i1,j1] = d[n1]
+           end
+       end
 
-        # loop over row
-        for i1 = 1 : L
-            idx = i1+i2-1
-            H[i1,i2] = v[idx]
-        end
+
+    # level 2
+    elseif N == 2
+       L1 = floor(Int64,dims[1]/2)+1; K1 = dims[1]-L1+1
+       L2 = floor(Int64,dims[2]/2)+1; K2 = dims[1]-L2+1
+
+       # allocate memory for Hankel matrix
+       H  = Array{Tv,2}(undef, L2*L1, K2*K1)
+
+       # build Hankel
+       for j2 = 1 : K2
+           c2 = (j2-1)*K1
+           for i2 = 1 : L2
+               r2 = (i2-1)*L1
+               n2 = i2 + j2 - 1
+
+               # first layer
+               for j1 = 1 : K1
+                   c1 = c2 + j1
+                   for i1 = 1 : L1
+                       r1 = r2 + i1
+                       n1 = i1 + j1 - 1
+                       H[r1, c1] = d[n1,n2]
+                   end
+               end #first
+           end
[…]

### 2019-07-05 src/agc.jl
arcis data denoising

@@ -1,0 +1,92 @@
+"""
+   compute the envelope of seismic data
+"""
+function envelope(d::Array{Tv}) where {Tv<:AbstractFloat}
+
+    # the dimensions of input
+    dims = size(d)
+
+    # matrix
+    if length(dims) == 2
+       dout = zeros(Tv, dims[1], dims[2])
+       for i = 1 : dims[2]
+           tmp = hilbert(d[:,i])
+           dout[:,i] .= abs.(tmp)
+       end
+
+    # cube
+    elseif length(dim) == 3
+       dout = zeros(Tv, dims[1], dims[2], dims[3])
+       for i3 = 1 : dims[3]
+           for i2 = 1 : dims[2]
+               tmp = hilbert(d[:,i2,i3])
+               dout[:,i2,i3] .= abs.(tmp)
+           end
+       end
+    end
+
+    return dout
+end
+
+"""
+   automatic gaining control, L1 is the number of windows for one trace,
+"""
+function agc(din::Array{Tv}; L1=10, L2=10) where {Tv<:AbstractFloat}
+
+    # get the dimensions
+    dims = size(din)
+    nt   = dims[1]
+    num_traces = prod(dims[2:end])
+
+    # reshape the input as matrix
+    din  = reshape(din, dims[1], num_traces)
+    dout = zeros(Tv, nt, num_traces)
+
+    # create smooth filter
+    hl        = floor(Int64, nt/L1)
+    op_length = 2 * hl + 1
+    smoother  = triang(op_length)
+    smoother  = convert(Vector{Tv}, smoother)
+
+    # envelope of the data
+    env  = envelope(din)
+    trip = floor(Int64, op_length/L2)
+
+    # loop over traces
+    for i = 1 : num_traces
+
+        # smoothing the envelope
+        tmp = env[:,i]
+        tmp = conv(tmp, smoother)
+
+        # truncate the result of smoothing to have same length
+        env_sm = tmp[hl+1 : nt+hl]
+        env_sm[1:trip]        .= env_sm[trip+1]
+        env_sm[nt-trip+1: end].= env_sm[nt-trip]
+
+        trace_in  = din[:, i]
+        trace_out = trace_in ./ env_sm
+        dout[:,i].= trace_out / maximum(abs.(trace_out))
+    end
+
+    return reshape(dout, dims...)
+end
+
+function balance(din::Array{Tv}) where {Tv<:AbstractFloat}
+
[…]

### 2019-07-05 src/patch/patch.jl

@@ -4,6 +4,11 @@ export get_patches_dir,
        taper_one_patch,
        par_taper,
        unpatch,
+       patch4d,
+       read_one_patch4d,
+       taper_one_patch4d,
+       par_taper4d,
+       unpatch4d,
        get_spatial_patches_dir,
        spatial_patch,
        read_one_spatial_patch,
@@ -12,4 +17,5 @@ export get_patches_dir,
        spatial_unpatch
 
 include("windowing.jl")
+include("windowing4d.jl")
 include("spatial_window.jl")

### 2019-07-05 src/patch/windowing4d.jl

@@ -1,0 +1,312 @@
+"""
+   get the directory of all patches, return a vector of String
+"""
+function get_patches_dir(path::String, it_nw::Ti, x1_nw::Ti, x2_nw::Ti, x3_nw::Ti) where {Ti<:Int64}
+
+    dir = Vector{String}(undef, it_nw * x1_nw * x2_nw * x3_nw)
+
+    for i3 = 1 : x3_nw
+        idx3 = (i3-1) * x2_nw * x1_nw * it_nw
+
+        for i2 = 1 : x2_nw
+            idx2 = idx3 + (i2-1) * x1_nw * it_nw
+
+            for i1 = 1 : x1_nw
+                idx1 = idx2 + (i1-1) * it_nw
+
+                for it = 1 : it_nw
+                    idx= idx1 + it
+
+                    file_name = join(["patch" "_" it "_" i1 "_" i2 "_" i3 ".bin"])
+                    dir[idx]  = joinpath(path, file_name)
+                end
+            end
+        end
+    end
+
+    return dir
+end
+
+"""
+   d is a 3D array, it_wl is the window length along time axis, it_wo is the
+overlapping of window.
+"""
+function patch4d(path::String, d::Array{Tv,4},
+               it_wl::Ti, it_wo::Ti,
+               x1_wl::Ti, x1_wo::Ti,
+               x2_wl::Ti, x2_wo::Ti,
+               x3_wl::Ti, x3_wo::Ti) where {Tv<:AbstractFloat, Ti<:Int64}
+
+    # size of data
+    (nt, n1, n2, n3) = size(d)
+
+    # shift in each direction
+    it_ws = it_wl - it_wo
+    x1_ws = x1_wl - x1_wo
+    x2_ws = x2_wl - x2_wo
+    x3_ws = x3_wl - x3_wo
+
+    # make sure the window size not too big
+    it_wl = it_wl > nt ? nt : it_wl
+    x1_wl = x1_wl > n1 ? n1 : x1_wl
+    x2_wl = x2_wl > n2 ? n2 : x2_wl
+    x3_wl = x3_wl > n3 ? n3 : x3_wl
+
+    # number of windows in each direction
+    it_nw = floor(Int64,(nt-it_wl)/it_ws)+1; it_nw = (it_nw-1)*it_ws+it_wl < nt ? it_nw+1 : it_nw
[…]

### 2019-07-05 src/tensor/tucker.jl

@@ -196,6 +196,61 @@ function local_tucker_als(cube::Array{Tv,3}, rk::Union{Vector{Ti},Ti}, work_dir:
      return s
 end
 
+### 4D tensor decomposition
+function local_tucker_als(cube::Array{Tv,4}, rk::Union{Vector{Ti},Ti}, work_dir::String;
+                          tol=1e-6, max_iter=10, init_flag="eigvec",
+                          it_wl = 30, it_wo = 10, x1_wl = 30, x1_wo = 10, x2_wl = 30, x2_wo = 10, x3_wl = 30, x3_wo = 10) where {Tv<:AbstractFloat, Ti<:Int64}
+
+     # divide the cube into patches
+     vec_dir = patch4d(work_dir, cube, it_wl, it_wo, x1_wl, x1_wo, x2_wl, x2_wo, x3_wl, x3_wo);
+
+     # pack arguments into a Dictionary
+     num_patches = length(vec_dir)
+     params = Vector{Dict}(undef, num_patches)
+     for i = 1 : num_patches
+         params[i] = Dict(:path=>vec_dir[i], :rk=>rk, :tol=>tol, :max_iter=>max_iter, :init_flag=>init_flag)
+     end
+
+     # wrap_fxy_prediction
+     function wrap_tucker_als(params::Dict)
+
+         # read one patch
+         (d, nt, n1, n2, n3, itl, itu, x1l, x1u, x2l, x2u, x3l, x3u, it_wo, x1_wo, x2_wo, x3_wo, code) = read_one_patch4d(params[:path], return_flag=true)
+
+         # denoising
+         s = tucker_als(Tensor(d), params[:rk]; tol=params[:tol], max_iter=params[:max_iter], init_flag=params[:init_flag])
+         s = tucker2tensor(s)
+
+         # write the result back
+         fid = open(params[:path], "r+")
+         pos = sizeof(Int64) * 17
+         seek(fid, pos)
+
+         if code == 1
+            write(fid, convert(Vector{Float64}, vec(s.d)))
+         elseif code == 2
+            write(fid, convert(Vector{Float32}, vec(s.d)))
+         end
+         close(fid)
+
+         return nothing
+     end
+
+     # apply fxy_prediction to each patch
+     pmap(wrap_tucker_als, params)
+
+     # taper the boundary of each patch
+     par_taper4d(vec_dir)
+
+     # merge patches
[…]

### 2019-07-05 src/mssa/fast_mssa.jl
testing Hankel multiplication

@@ -468,6 +468,321 @@ end
 
 
 # Hankel matrix or its Hermitian transpose times a vector
+function hankel_multiplication(d::Array{Tv}, v::Array{Tv}; flag="forward") where {Tv<:Number}
+
+    # dimensions of array
+    dims = size(d)
+    N    = length(dims)
+
+    # Fourier transform of d
+    d_hat = fft(d)
+
+    if N == 1
+       # compute L, K
+       L1 = floor(Int64,dims[1]/2)+1; K1 = dims[1]-L1+1
+
+       if flag == "forward"
+
+          # check the size of v
+          size(v, 1) == K1 ||  DimensionMismatch()
+
+          # reverse v and padding zeros
+          v_hat = reverse_order(v; n1=dims[1])
+          v_hat = fft(v_hat)
+
+          # element-wise multiplication
+          r = ifft(d_hat .* v_hat)
+
+          # return the last L1 element
+          return r[K1:dims[1]]
+
+       elseif flag == "adjoint"
+
+          # check the size of v
+          size(v, 1) == L1 ||  DimensionMismatch()
+
+          # reverse v and padding zeros
+          v_hat = reverse_order(v; n1=dims[1])
+          v_hat = fft(v_hat)
+
+          # conjugate property
+          d_tilde = copy(d_hat)
+          for i1 = 1 : dims[1]
+              j1 = i1 == 1 ? 1 : dims[1]-i1+2
+
+              d_tilde[i1] = conj(d_hat[j1])
+          end
+
+          # element-wise multiplication
+          r = ifft(d_tilde .* v_hat)
+
+          # return the last K1 element
+          return r[L1:dims[1]]
+
+       else
+          error("non-surpported operation")
+       end
+
+
+    elseif N == 2
+       # compute L, K
+       L1 = floor(Int64,dims[1]/2)+1; K1 = dims[1]-L1+1
+       L2 = floor(Int64,dims[2]/2)+1; K2 = dims[2]-L2+1
+
+       if flag == "forward"
+
+          # check the size of v
+          size(v, 1) == K1 || DimensionMismatch()
+          size(v, 2) == K2 || DimensionMismatch()
+
+          # reverse v and padding zeros
+          v_hat = reverse_order(v; n1=dims[1], n2=dims[2])
+          v_hat = fft(v_hat)
+
[…]

### 2020-01-28 doc/fast_ssa/paper.tex
wirtinger derivative

@@
 $$
-{\mathbfcal F}[x(t-t_0)] = e^{i \omega t_0} \int_{-\infty}^{\infty} x(t^{\prime}) e^{-i\omega t^{\prime}} d t^{\prime} = X(\omega) e^{i \omega t_0}
+{\mathbfcal F}[x(t-t_0)] = e^{-i \omega t_0} \int_{-\infty}^{\infty} x(t^{\prime}) e^{-i\omega t^{\prime}} d t^{\prime} = X(\omega) e^{-i \omega t_0}
 $$
@@
 
+\section{Complex derivative}
+Let $f(z, z^{\ast})$: $\mathbb{C} \times \mathbb{C} \rightarrow \mathbb{C}$ be a function of a complex variable $z$ and its complex conjugate $z^{\ast}$, where $z = x + iy$. We have the derivative 
+\begin{equation}
+\frac{\partial f}{\partial z} = \frac{1}{2}(\frac{\partial f}{\partial x} - i\frac{\partial f}{\partial x}),
+\end{equation}
+and
+\begin{equation}
+\frac{\partial f}{\partial z^{\ast}} = \frac{1}{2}(\frac{\partial f}{\partial x} + i\frac{\partial f}{\partial x}).
+\end{equation}
+Using the definition of complex derivative, it's clear that 
+\begin{equation}
+\begin{split}
+\frac{\partial f}{\partial z} &= \left( \frac{\partial f}{\partial z^{\ast}} \right)^{\ast} \\
+\frac{\partial z}{\partial z} &= 1 \\
+\frac{\partial z^{\ast}}{\partial z^{\ast}} &= 1 \\
+\frac{\partial z^{\ast}}{\partial z} &= 0
+\end{split}.
+\end{equation}
+With above property, we can simply consider the complex variable and its complex conjugate $z^{\ast}$ as two independent variable.
+
+\subsection{Stationary point}
+Let $f(z, z^{\ast})$: $\mathbb{C} \times \mathbb{C} \rightarrow \mathbb{R}$ be a function mapping a complex variable $z$ and its complex conjugate $z^{\ast}$ to a real number.
+A sufficient and necessary condition for $f(z, z^{\ast})$ have a stationary point is $\frac{\partial f}{\partial z} = 0$.
+Similarly $\frac{\partial f}{\partial z^{\ast}} = 0$ is also a sufficient and necessary condition.
+
+\subsection{Complex gradient}
+Let's think about the first-order Taylor-series expansion, we have 
+\begin{equation}
+\delta f = \frac{\partial f}{\partial z} \delta z +  \frac{\partial f}{\partial z^{\ast}} \delta z^{\ast} 
+\end{equation}
+By using the property $\frac{\partial f}{\partial z} = \left( \frac{\partial f}{\partial z^{\ast}} \right)^{\ast}$, we can get 
+\begin{equation}
+\delta f = 2 Re (\frac{\partial f}{\partial z} \delta z),
+\end{equation}
+where the operator $Re$ means take the real part of the input.
+We can extend this property to a function of a complex vector $g$: $\mathbb{C}^N \rightarrow \mathbb{R}$.
+We denote the complex vector as $\bf z$, so its first-order Taylor expansion can be expressed as
+\begin{equation}
+\begin{split}
+\delta g &= 2 \sum_{i=1}^N Re \left(\frac{\partial g}{\partial z_i} \delta z_i \right) \\
+             &= 2 Re\left( \left( \frac{\partial g}{\partial {\bf z}} \right)^T {\bf \delta z} \right)
+\end{split}
+\end{equation}
+To make the above expression consistent with the axiom of inner product, we replace the above equation as 
+\begin{equation}
+\delta g = 2 Re \left( \left( \frac{\partial g}{\partial {\bf z^{\ast}}}\right)^H {\bf \delta z} \right)
+\end{equation}
+The maximum change $\delta g$ is obtained when $\delta {\bf z}$ is in the same direction of $\frac{\partial g}{\partial {\bf z^{\ast}}}$. It means that
+\begin{equation}
+\delta {\bf z} = \alpha \frac{\partial g}{\partial {\bf z^{\ast}}}.
+\end{equation}
+Where $\alpha$ is a real number indicate the magnitude of the vector $\delta {\bf z}$.
+This explanation suggest that the gradient vector $ \frac{\partial g}{\partial {\bf z^{\ast}}}$ is required for updating the variable ${\bf z}$ in an optimization problem.
+
+\subsection{discrete time linear time invariant system}
+We derive the mathematical model for describing discrete time linear invariant system.
+Suppose the input to a system is an impulse $\delta[n]$, which is equal to $1$ when $n=0$ and is $0$ elsewhere.
+The impulse response of this system is indicated by $h[n]$.
+The $\textbf{time invariant}$ property mean that the delayed impulse response is the output of the delayed impulse, which can be expressed as
+\begin{equation}
+\delta[n-k] \rightarrow h[n-k],
+\end{equation}
+where $k$ indicate the number of samples delayed.
+
+Any discrete-time signal(sequence) can be expressed as the weighted summation of delayed impulse as 
+\begin{equation}
+x[n] = \sum_{k=-\infty}^{+\infty} x[k] \delta[n-k]
+\label{input}
+\end{equation}
+Combine with the linear property of this system, the output can be expressed as
+\begin{equation}
+y[n] = \sum_{k=-\infty}^{+\infty} x[k] h[n-k]
+\end{equation}
+Which give the definition of convolution which are used to describe the linear time-invariant system as $y[n] = x[n] \ast h[n]$.
+
+Suppose the input is given as $g[n] = x[n-p_1]$ and the impulse response of the system is $f[n] = h[n-p_2]$.
+What is the output of this system?
+There are two ways to derive the result, one is based on the definition we given in equation (\ref{input}).
+\begin{equation}
+\begin{split}
+g[n] &= \sum_{k=-\infty}^{+\infty} g[k] \delta[n-k] \\
+       &= \sum_{k=-\infty}^{+\infty} x[k-p_1] \delta[n-k]
+\end{split}
+\end{equation}
+The output of this system is
+\begin{equation}
+\begin{split}
+s[n] &= \sum_{k=-\infty}^{+\infty} x[k-p_1] f[n-k] \\
+       &= \sum_{k=-\infty}^{+\infty} x[k-p_1] h[n-k-p_2] \\
+\end{split}
+\label{sum}
+\end{equation}
+We can the variable change trick by setting $k^{\prime} = k-p_1$, so we have $k = k^{\prime} + p_1$. With this variable change, equation (\ref{sum}) can be reformulated as 
+\begin{equation}
+\begin{split}
+s[n]  &= \sum_{k^{\prime}=-\infty}^{+\infty} x[k^{\prime}] h[n-k^{\prime}-p_1- p_2]    \\
+        &= \sum_{k^{\prime}=-\infty}^{+\infty} x[k^{\prime}] h[(n-p_1- p_2)-k^{\prime}] \\
+        &= y[n-p_1-p_2]
+\end{split}
+\label{sum}
+\end{equation}
+
 \bibliographystyle{seg}

### 2020-08-12 doc/fast_ssa/paper.tex
sync the content

@@
 The above operation can be expressed into the matrix-vector form.
-To simplify the definition of Fourier transform matrix, we define a variable $z = e^{-i \pi  / N}$, so the Fourier transform matrix, which is a $N$-by-$N$ square matrix and the element $F_{kl} = z^{kl}$.
+To simplify the definition of Fourier transform matrix, we define a variable $z = e^{-i 2\pi  / N}$, so the Fourier transform matrix, which is a $N$-by-$N$ square matrix and the element $F_{kl} = z^{kl}$.
 It can be expanded as
 \begin{equation}
-{\bf F}_N = \begin{bmatrix}
+{\bf F}_N = \frac{1}{\sqrt{N}}\begin{bmatrix}
 1 & 1 & \cdots & 1 \\
@@
 \begin{equation}
-{\bf F}_N^{-1} = \frac{1}{N} {\bf F}_N^H = \frac{1}{N} {\bf F}_N^{\ast} \,
+{\bf F}_N^{-1} = {\bf F}_N^H =  {\bf F}_N^{\ast} \,
 \label{eq6}
 \end{equation}
-where the super-script $^H$ represent complex conjugate transpose and $\ast$ indicate conjugate transpose. As the matrix ${\bf F}_N$ is a symmetrical matrix, that's why we can get equation \ref{eq6}.
+where the super-script $^H$ represent complex conjugate transpose and $\ast$ indicate complex conjugate to each element.
+As the matrix ${\bf F}_N$ is a symmetrical matrix, that's why we can get equation \ref{eq6}.
 
@@
 
-\subsection{Circulant matrix}
+\subsection{   matrix}
 An $9 \times 9$ circulant matrix is defined as
@@
 \begin{equation}
-{\bf Cv} = {\bf F}_N^{-1} \left(  {\bf F}_N {\bf c} \circ {\bf F}_N {\bf v} \right)  =  {\bf F}_N^{-1} \left(  diag({\bf F}_N {\bf c}) \cdot {\bf F}_N {\bf v} \right) \,,
+{\bf Cv} = {\bf F}_N^{-1} \left(  \sqrt{N}{\bf F}_N {\bf c} \circ {\bf F}_N {\bf v} \right)  =  {\bf F}_N^{-1} \left( \sqrt{N} diag({\bf F}_N {\bf c}) \cdot {\bf F}_N {\bf v} \right) \,,
+\label{circ_times_vector}
 \end{equation}
 where the symbol $\circ$ and $diag$ represents Hardmard (element-wise) multiplication and building diagonal matrix from a vector.
+Here we use a $4$-point vector to explain the above property, we define a vector ${\bf c} = \begin{bmatrix} c_1 & c_2 &c_3 & c_4\end{bmatrix}^T$. The circulant matrix formed by this vector is given as
+\begin{equation}
+{\bf C} = 
+\begin{bmatrix}
+c_1 & c_4 & c_3 & c_2  \\ 
+c_2 & c_1 & c_4 & c_3  \\ 
+c_3 & c_2 & c_1 & c_4  \\ 
+c_4 & c_3 & c_2 & c_1 
+\end{bmatrix}
+\end{equation} 
+Follow above definitions, we give $z = e^{-i 2\pi / 4}$, so the corresponding Fourier transform matrix is given as
+\begin{equation}
+{\bf F}_4 = \frac{1}{2}\begin{bmatrix}
+1 & 1 & 1 & 1 \\
+1 & z & z^2 & z^3 \\
+1 & z^2 & z^4 & z^6 \\
+1 & z^3 & z^6 & z^9 \\
+\end{bmatrix}
+\end{equation}
+The inverse Fourier transform matrix is the complex conjugate transpose as
+\begin{equation}
+{\bf F}_4^H = \frac{1}{2}\begin{bmatrix}
+1 & 1 & 1 & 1 \\
+1 & z^{-1} & z^{-2}  & z^{-3}  \\
+1 & z^{-2}  & z^{-4}  & z^{-6}  \\
+1 & z^{-3}  & z^{-6}  & z^{-9}  \\
+\end{bmatrix}
+\end{equation}
+The multiplication between the circulant matrix and the first column of ${\bf F}_4^H$ is given as
+\begin{equation}
+{\bf C}{\bf f}_1 = \frac{1}{2}
+\begin{bmatrix}
+c_1 & c_4 & c_3 & c_2  \\ 
+c_2 & c_1 & c_4 & c_3  \\ 
+c_3 & c_2 & c_1 & c_4  \\ 
+c_4 & c_3 & c_2 & c_1 
+\end{bmatrix}
+\begin{bmatrix}
+1 \\
+1 \\
+1 \\
+1 
+\end{bmatrix}
+= \frac{c_1+c_2+c_3+c_4}{2}
+\begin{bmatrix}
+1 \\
+1 \\
+1 \\
+1 
+\end{bmatrix}
+\end{equation} 
+The coefficient in front of the vector is equation to the first component of the Fourier transform of vector ${\bf c}$ times $\sqrt{N}$, where $N$ is the total length of the vector.
+Let's continue the discussion and multiply the circulant matrix by the second column of ${\bf F}_4^H$. We get
+\begin{equation}
+{\bf C}{\bf f}_1 = \frac{1}{2}
+\begin{bmatrix}
+c_1 & c_4 & c_3 & c_2  \\ 
+c_2 & c_1 & c_4 & c_3  \\ 
+c_3 & c_2 & c_1 & c_4  \\ 
+c_4 & c_3 & c_2 & c_1 
+\end{bmatrix}
+\begin{bmatrix}
+1 \\
+z^{-1} \\
+z^{-2} \\
+z^{-3} 
+\end{bmatrix}
+= \frac{1}{2}
+\begin{bmatrix}
+c_1 + c_4 z^{-1} + c_3 z^{-2} + c_2 z^{-3}  \\
+c_2 + c_1 z^{-1} + c_4 z^{-2} + c_3 z^{-3}  \\
+c_3 + c_2 z^{-1} + c_1 z^{-2} + c_4 z^{-3}  \\
+c_4 + c_3 z^{-1} + c_2 z^{-2} + c_1 z^{-3}  \\
+\end{bmatrix}
+= \frac{1}{2}
+\begin{bmatrix}
+1\left(c_1 + c_4 z^{-1} + c_3 z^{-2} + c_2 z^{-3} \right)  \\
+z^{-1} \left(c_2 z    + c_1 + c_4 z^{-1} + c_3 z^{-2} \right)  \\
+z^{-2} \left(c_3 z^2+ c_2 z + c_1 + c_4 z^{-1} \right)  \\
+z^{-3} \left(c_4 z^3+ c_3 z^{-2} + c_2 z + c_1 \right) \\
+\end{bmatrix}
+\end{equation}
+The function $z^n$ is a periodic function with period of $N=4$, the vector on the right hand side is equal to
+\begin{equation}
+\frac{1}{2}
+\begin{bmatrix}
+1\left(c_1 + c_4 z^{-1} + c_3 z^{-2} + c_2 z^{-3} \right)  \\
+z^{-1} \left(c_2 z    + c_1 + c_4 z^{-1} + c_3 z^{-2} \right)  \\
+z^{-2} \left(c_3 z^2+ c_2 z + c_1 + c_4 z^{-1} \right)  \\
+z^{-3} \left(c_4 z^3+ c_3 z^{-2} + c_2 z + c_1 \right) \\
+\end{bmatrix}
+=
+\frac{1}{2}
+\begin{bmatrix}
+1         \left(c_1 + c_2 z + c_3 z^2 + c_4 z^3 \right)  \\
+z^{-1} \left(c_1 + c_2 z + c_3 z^2 + c_4 z^3 \right)  \\
+z^{-2} \left(c_1 + c_2 z + c_3 z^2 + c_4 z^3 \right)  \\
+z^{-3} \left(c_1 + c_2 z + c_3 z^2 + c_4 z^3 \right) \\
+\end{bmatrix} 
+= 
+\frac{1}{2}\left(c_1 + c_2 z + c_3 z^2 + c_4 z^3 \right) 
+\begin{bmatrix}
+1         \\
+z^{-1} \\
+z^{-2} \\
+z^{-3} \\
+\end{bmatrix} 
+\end{equation}
+The define the scalar in front of the right vector as $\lambda = \frac{1}{2}\left(c_1 + c_2 z + c_3 z^2 + c_4 z^3 \right) $, which is the second element of the Fourier transform of the vector $c$ times $\sqrt{N}$.
+Following the same derivation, we can get
+\begin{equation}
+{\bf C} {\bf f}_i = \lambda_i  {\bf f}_i
+\end{equation}
+If we organize all the vectors into a matrix, we can get
+\begin{equation}
+{\bf C}{\bf F}^H = {\bf F}^H diag(\sqrt{N}{\bf F}{\bf c})
+\end{equation}
+We can further right multiply both end of the equations with ${\bf F}$, we can get
+\begin{equation}
+{\bf C} = {\bf F}^H diag(\sqrt{N}{\bf F}{\bf c}) {\bf F}
+\label{decom}
+\end{equation}
+Which is equivalent to the expression in equation \ref{circ_times_vector}.
+In reality, we never formulate the Fourier transform matrix explicitly, instead the fast Fourier transform operator $fft$ and the inverse transform operator $ifft$ are used.
+we know
+\begin{equation}
+\begin{split}
+\sqrt{N} {\bf F}_n = fft \\
+ {\bf F}_N^H = \sqrt{N} ifft
+ \end{split}
+\end{equation}
+with the above expression, we can replace the equation \ref{circ_times_vector} with the two operator as
+\begin{equation}
+{\bf C} {\bf v} = \sqrt{N} ifft\left( diag(fft({\bf c}))  {\frac{1}{\sqrt{N}} fft({\bf v})} \right)
+\label{decom}
+\end{equation}
+With some simple manipulations, we can get the final form as
+\begin{equation}
+{\bf C} {\bf v} =  ifft \left( fft({\bf c}) \circ  fft({\bf v}) \right)
+\label{decom}
+\end{equation}
 

### 2020-08-12 examples/function_creator.jl

@@ -1,0 +1,108 @@
+"""
+   return two functions
+"""
+function hankel_multiplication_creator(d::Array{Tv}) where {Tv <: Union{Float32, Float64, Complex{Float32}, Complex{Float64}}}
+
+    # dimensions of array
+    dims = size(d)
+    N    = length(dims)
+
+    # Fourier transform of d
+    d_hat = fft(d)
+
+    # compute L, K
+    L1 = floor(Int64,dims[1]/2)+1; K1 = dims[1]-L1+1
+    L2 = floor(Int64,dims[2]/2)+1; K2 = dims[2]-L2+1
+    L3 = floor(Int64,dims[3]/2)+1; K3 = dims[3]-L3+1
+    L4 = floor(Int64,dims[4]/2)+1; K4 = dims[4]-L4+1
+
+    # create forward operator
+    op_forward = function (x::Array{Tv})
+
+                     # dimension of x
+                     M = ndims(x)
+
+                     # check the size of input
+                     K4*K3*K2*K1 == length(x) || throw(DimensionMismatch("check length of input"))
+
+                     # reshape x to a multi-dimensonal array, reshape doesn't change the input
+                     x = reshape(v, K1, K2, K3, K4)
+
+                     # reverse x and padding zeros, new memory is allocated
+                     x_hat = reverse_order(x; n1=dims[1], n2=dims[2], n3=dims[3], n4=dims[4])
+                     fft!(x_hat)
+
+                     # element-wise multiplication
+                     x_hat .= x_hat .* d_hat
+                     ifft!(x_hat)
+
+                     # make the output have same data format
+                     if Tv <: AbstractFloat
+
+                        # make the output have same dimension
+                        if M == 1
+                           return vec(real(x_hat[K1:dims[1],K2:dims[2],K3:dims[3],K4:dims[4]]))
+                        else
+                           return real(x_hat[K1:dims[1],K2:dims[2],K3:dims[3],K4:dims[4]])
+                        end
+
+                     else
+
+                        # make the output have same dimension
+                        if M == 1
+                           return vec(x_hat[K1:dims[1],K2:dims[2],K3:dims[3],K4:dims[4]])
+                        else
+                           return x_hat[K1:dims[1],K2:dims[2],K3:dims[3],K4:dims[4]]
+                        end
+                     end
+                 end
+
+    # create adjoint operator
+    op_adjoint = function (x::Array{Tv})
+
+                     # dimension of x
+                     M = ndims(x)
+
+                     # check the size of input
[…]

### 2020-08-12 examples/linear_events.jl

@@ -5,37 +5,30 @@ dt = 4.0/1000.
 # generate data with two linear events
 d = SeisLinearEvents(;ot=0.0,dt=dt, nt=100, ox1=0.0, dx1=10.0,
                       nx1=40, ox2=0.0, dx2=10.0, nx2=40, ox3=0.0, dx3=10.0,
-                      nx3=1, ox4=0.0, dx4=10.0, nx4=1, tau=[.1,.25],
-                      p1=[0.0001,-0.0003],p2=[0.0001,-0.0003],p3=[0.,0],p4=[0.,0.],
-                      amp=[1.0,-1.0], f0=24.0);
+                      nx3=1, ox4=0.0, dx4=10.0, nx4=1, tau=[.1,.25,0.3],
+                      p1=[0.0001,-0.0003,0.0002],p2=[0.0001,-0.0003,0.0002],p3=[0.,0.,0.],p4=[0.,0.,0.],
+                      amp=[1.0,-1.0,1.5], f0=24.0);
 
 # add random noise to data
 dn = SeisAddNoise(d, 0.5);
 
-# tucker decomposition
-s1 = tucker_als(Tensor(dn), 12; init_flag="eigvec");
-s1 = tucker2tensor(s1).d;
+# fxy eigen
+s1 = fxy_eigen(dn, 3; dt=dt, flow=0.0, fhigh=60.0);
 
 # MSSA
-s2 = mssa(dn, 2; dt=dt, flow=0.0, fhigh=60.0);
+s2 = mssa(dn, 3; dt=dt, flow=0.0, fhigh=60.0);
 
 # fxy prediction
-s3 = fxy_prediction(dn, 2; dt=dt, flow=2.0, fhigh=60.0,
+s3 = fxy_prediction(dn, 3; dt=dt, flow=2.0, fhigh=60.0,
                     max_iter=10, mu=1.0e-6, tol=1.0e-9);
-
-# fxy eigen
-s4 = fxy_eigen(dn, 2; dt=dt, flow=0.0, fhigh=60.0);
 
 # plotting the result
 i = 10;
 tmp1 = hcat(s1[:,:,i], dn[:,:,i]-s1[:,:,i]);
 tmp2 = hcat(s2[:,:,i], dn[:,:,i]-s2[:,:,i]);
 tmp3 = hcat(s3[:,:,i], dn[:,:,i]-s3[:,:,i]);
[…]

### 2020-08-12 examples/mixture_events.jl

@@ -1,8 +1,10 @@
 using SeisPlot, SeisDenoise, SeisProcessing
 
+# time sample
 dt = 8.0/1000.
 
-d = get_mixture_events(nt=250, nx1=200, nx2=200, dt=dt, dx1=10, dx2=10,
+# synthetic data
+d = get_mixture_events(nt=250, nx1=120, nx2=120, dt=dt, dx1=10, dx2=10,
                               v1=[6000. ,2000,-2700.,-2700.,6000. ],
                               v2=[14000.,4000, 4000., 4000.,14000.],
                               tau=[0.1  ,0.4 ,  0.9 ,  1.2 ,1.4   ],
@@ -10,20 +12,22 @@ d = get_mixture_events(nt=250, nx1=200, nx2=200, dt=dt, dx1=10, dx2=10,
                               apx1=[0.0 ,0.0 ,  0.0 ,  0.0 ,0.0   ],
                               apx2=[0.0 ,0.0 ,  0.0 ,  0.0 ,0.0   ],
                               event_type=['l', 'h', 'p', 'p', 'l' ],
-                              f0=20.0, static_flag=true, L=13, sigma=0.1);
+                              f0=20.0, static_flag=true, L=13, sigma=0.03);
 
 # add random noise to data
 dn = SeisAddNoise(d, 0.75);
 
 # work directory
-work_dir = joinpath(homedir(), "Desktop/randCP");
+work_dir = joinpath(homedir(), "Desktop/patches");
+isdir(work_dir) && rm(work_dir, recursive=true);
+mkdir(work_dir);
 
 # fxy eigen
 s1 = local_fxy_eigen(dn, 4, work_dir; dt=dt, flow=2.0, fhigh=60.0,
                      x1_wl = 30, x1_wo = 10, x2_wl = 30, x2_wo = 10);
 
 # mssa
-s2 = local_mssa(dn, 18, work_dir; dt=dt, flow=2.0, fhigh=60.0,
+s2 = local_mssa(dn, 2, work_dir; dt=dt, flow=2.0, fhigh=60.0,
[…]

### 2020-08-12 src/deblending/SLO.jl

@@ -1,0 +1,100 @@
+function apply_time_shift(d::Matrix{Tv}, dither::Vector, dt) where {Tv<:Real}
+
+    # to eliminate the influence of periodic property of DFT
+    # maybe worth to padding zeros to both ends, no implemented at here.
+
+    # get dimension of data
+    (n1, n2) = size(d)
+    n2 == length(dither) || error("check the length of dither")
+
+    # fourier transform along time axis
+    D = fft(d, 1)
+
+    # radian frequency interval
+    delta_w = 2 * pi * (1.0 / dt / n1)
+
+    # compute half of the radian frequency axis
+    # if n1 is even, nyquist frequency is computed, otherwise not
+    nw = floor(Int64, n1/2) + 1
+    w  = zeros(Tv, nw)
+    for i = 1 : nw
+        w[i] = (i-1) * delta_w
+    end
+
+    # apply time shift to each trace
+    for i2 = 1 : n2
+        for i1 = 2 : nw  # start from first non-zero frequency component
+
+            # the conjugate property of real series
+            j1 = n1 - i1 + 2
+            D[i1,i2] = D[i1,i2] * exp(-im * w[i1] * dither[i2])
+            D[j1,i2] = conj(D[i1,i2])
+        end
+    end
+
+    # transform back to time domain
+    return real(ifft(D,1))
+
+end
+
+"""
+   From second dimension determine the number of boat
+"""
+function deblending_by_inversion(d::Matrix{Tv}, dither::Array{Tv1}, trace_length, dt;
+         max_iter::Ti=100, alpha=0.9, truncate=[]) where {Tv<:AbstractFloat, Ti<:Int64, Tv1<:Real}
+
+     # number of simultaneous source
+     num_src = size(dither, 2)
+
+     # get the dimension of super gather
+     (n1, n2) = size(d)
+
+     # recording length (number of samples)
+     ns = floor(Int64, trace_length/dt) + 1
+
+     # determine the thresholding curve
+     threshold = zeros(Complex{Float64}, max_iter, num_src)
+     for i = 1 : num_src  # loop over boat
+
+         # aline traces for current source
+         if  norm(dither[:,i]) > 0.0
+
+             ds = apply_time_shift(d, -dither[:,i], dt) # to aline the signal, need to shift in opposite direction
+
[…]

### 2020-08-12 src/deblending/time_shift.jl

@@ -1,0 +1,136 @@
+# ==============================================================================
+function time_shift(d::Matrix{Tv}, dither::Vector, dt) where {Tv<:Real}
+
+    # to eliminate the influence of periodic property of DFT
+    # maybe worth to padding zeros to both ends, no implemented at here.
+
+    # get dimension of data
+    (n1, n2) = size(d)
+    n2 == length(dither) || error("check the length of dither")
+
+    # fourier transform along time axis
+    D = fft(d, 1)
+
+    # radian frequency interval
+    delta_w = 2 * pi * (1.0 / dt / n1)
+
+    # compute half of the radian frequency axis
+    # if n1 is even, nyquist frequency is computed, otherwise not
+    nw = floor(Int64, n1/2) + 1
+    w  = zeros(Tv, nw)
+    for i = 1 : nw
+        w[i] = (i-1) * delta_w
+    end
+
+    # apply time shift to each trace
+    for i2 = 1 : n2
+        for i1 = 2 : nw  # start from first non-zero frequency component
+
+            # the conjugate property of real series
+            j1 = n1 - i1 + 2
+            D[i1,i2] = D[i1,i2] * exp(-im * w[i1] * dither[i2])
+            D[j1,i2] = conj(D[i1,i2])
+        end
+    end
+
+    # transform back to time domain
+    return real(ifft(D,1))
+
+end
+
+function blend_forward(d1, d2, time_delay, dt)
+    # apply time delay to far boat
+    d3 = time_shift(d2, time_delay, dt)
+    return d1 + d3
+end
+
+function blend_adjoint(d_blend, time_delay, dt)
+
+    d1 = copy(d_blend)
+    d2 = time_shift(d_blend, -time_delay, dt)
+
+    return d1, d2
+end
+
+function hard_threshold!(d::Matrix{Complex{Tv}}, gamma::Tv) where {Tv<:AbstractFloat}
+
+    (n1, n2) = size(d)
+    for i2 = 1 : n2
+        for i1 = 1 : n1
+            if abs(d[i1,i2]) < gamma
+               d[i1,i2] = 0.0im
+            end
+        end
+    end
+    return nothing
+end
+
+function deblending_by_inversion(d_blend::Matrix{Tv}, time_delay::Vector{Tv}, dt::Tv;
+         max_iter=200, min_value=1e-5) where {Tv<:AbstractFloat}
+
[…]

### 2020-08-12 src/median_filter/scaler2d.jl

@@ -1,0 +1,64 @@
+"""
+   h is the window half-length in vertical direction and w is the half-length in
+horizontal direction.
+"""
+function scalar_median_filter(d::Matrix{Tv}, h::Ti, w::Ti) where {Tv<:Real, Ti<:Int64}
+
+    (n1, n2) = size(d)
+    d_filter = zeros(Tv, n1, n2)
+
+    for i2 = 1 : n2
+        for i1 = 1 : n1
+
+            # vertical range
+            i1_lower = i1 - h < 1  ? 1  : i1 - h
+            i1_upper = i1 + h > n1 ? n1 : i1 + h
+
+            # horizontal range
+            i2_lower = i2 - w < 1  ? 1  : i2 - w
+            i2_upper = i2 + w > n2 ? n2 : i2 + w
+
+            v = median(view(d, i1_lower:i1_upper, i2_lower:i2_upper))
+            d_filter[i1,i2] = v
+        end
+    end
+    return d_filter
+end
+
+function vector_median_filter(d::Matrix{Tv}, h::Ti, w::Ti) where {Tv<:Real, Ti<:Int64}
+
+
+
+end
+
+
+
+dcut = d[281:449,20,77:128];
+SeisPlotTX(dcut, style="wiggles", xcur=3.0);
+
+d1 = scalar_median_filter(dcut, 2, 2);
+SeisPlotTX(d1, style="wiggles", xcur=3.0);
+
+
+dir_work = joinpath(homedir(), "Desktop/PGS/PGS_rsf");
+
+path_d   = joinpath(dir_work, "data.rsf");
+(hdr, d) = read_RSdata(path_d);
+for i = 40 : 40 : size(d,2)
+    SeisPlotTX(d[:,i,:], hbox=8, wbox=5, pclip=90, cmap="gray"); tight_layout();
+end
+
+idx_channel = 100; dt = 0.016;
+
+d1 = scalar_median_filter(d[:,idx_channel,:], 1, 1);
+
+i1l = 1; i1u = size(d,1); ot = (i1l-1)*dt; tmax = (i1u-1)*dt; i1_range = range(ot, length=5, stop=tmax);
+i2l = 1; i2u = size(d,3); i2_range = range(i2l, length=5, stop=i2u);
+SeisPlotTX(d[i1l:i1u,idx_channel,i2l:i2u], hbox=8, wbox=6, pclip=95, cmap="gray",
[…]

### 2020-08-12 src/mssa/lanczos.jl

@@ -1,0 +1,791 @@
+"""
+   Lanczos bidiagonalization with partial reorthogonalization, A can be matrix or
+forward operator. k is the steps. r0 is the starting vector. the output U is a m*k
+matrix, Bk is a k*k bidiagonal matrix and V is a n*k matrix and the residue vector
+r0 make A*V_{k} = U_{k} * B_{k} + u_{k+1} * beta_{k+1} * e_{k}' and r0 = u_{k+1} * beta[k+1]
+
+the call of the functon is
+(U, Bk, V, r0, ierr, workload) = lanbpro(A, k, r0)
+
+workload =[numReoU, numIptU; numReoV, numIptV]
+
+numReoU: number of orthgonalization for U
+numIptU: number of inner product for U
+numReoV: number of orthgonalization for V
+numIptV: number of inner product for V
+
+"""
+function lanbpro(A::Union{Matrix{Tv}, Function}, k::Ti; m=[], n=[], params=[],
+                 delta=[], Anorm=[], Uold=[], Bold=[], Vold=[], r0=[],
+                 examine_flag=false, elr_flag=true) where {Ti<:Int64, Tv <: Union{Float32, Float64, Complex{Float32}, Complex{Float64}}}
+
+    # decide A is an operator or matrix
+    matrix_flag = isa(A, Matrix)
+
+    # A is a matrix
+    if matrix_flag
+       etype  = eltype(A)
+       (m, n) = size(A)
+
+    # A is function
+    else
+       if m == [] || n==[]
+          error("must specify the size of operator A")
+       end
+       etype = Float64
+    end
+
+    # numerical error
+    smleps = eps(etype)
+
+    # relaxiation coefficient for the inner product after reorthogonalization
+    m2 = etype(3.0 / 2.0)
+    n2 = etype(3.0 / 2.0)
+
+    # desired level of orthogonality (the defult value)
+    delta = delta == [] ? sqrt(smleps/k) : delta
+
+    # critia for selecting subset of vectors
+    eta   = etype(smleps^(3.0 / 4.0) / sqrt(k))
+
+    # the estimated error of matrix multiplication
+    eps1  = etype(sqrt(maximum([m n])) * eps(etype) / 2.0)
+
+    # force orthogonalization at every step (still select subset of vectors to orthogonal aginst)
+    forceRO_flag = delta == 0 ? true : false
+
+    # parameter for Gram-schmit method
[…]

### 2020-08-12 src/mssa/mssa.jl

@@ -10,8 +10,11 @@ export hankelization,
        reverse_order,
        hankel_multiplication,
        count_copy_times,
-       anti_diagonal_summation
+       anti_diagonal_summation,
+       anti_diagonal_summation_slow,
+       lanbpro
 
 include("mssa3d.jl")
+include("lanczos.jl")
 include("fxy_eigen.jl")
 include("structured_matrix.jl")

### 2020-08-12 src/mssa/structured_matrix.jl

@@ -1,7 +1,7 @@
 """
    build a multi-level circulant matrix up to five dimension
 """
-function build_circulant_matrix(d::Array{Tv}) where {Tv <: Union{Float32, Float64, Complex{Float32}, Complex{Float64}}}
+function build_circulant_matrix(d::Array{Tv}) where {Tv <: Number}
 
     # determine the dimensions of an array
     dims = size(d)
@@ -195,7 +195,7 @@ end
 """
    build a multi-level Toeplitze matrix from a multi-dimensional array
 """
-function build_toeplitz_matrix(d::Array{Tv}) where {Tv <: Union{Float32, Float64, Complex{Float32}, Complex{Float64}}}
+function build_toeplitz_matrix(d::Array{Tv}) where {Tv <: Number}
 
     # determine the dimensions of an array
     dims = size(d)
@@ -404,7 +404,7 @@ end
 """
    build a multi-level block Hankel matrix from a multi-dimensional array
 """
-function build_hankel_matrix(d::Array{Tv}) where {Tv <: Union{Float32, Float64, Complex{Float32}, Complex{Float64}}}
+function build_hankel_matrix(d::Array{Tv}) where {Tv <: Number}
 
     # determine the dimensions of an array
     dims = size(d)
@@ -610,11 +610,18 @@ end
 to the resultant multi-dimensional array
 """
 function reverse_order(d::Array{Tv}; n1=0, n2=0, n3=0, n4=0, n5=0
-         ) where {Tv <: Union{Float32, Float64, Complex{Float32}, Complex{Float64}}}
+         )  where {Tv <: Number}
 
     # get the dimensions of input array
     dims = size(d)
     N    = length(dims)
+
+    # determine single or double precision
+    if Tv <: Union{Int32, Float32, Complex{Float32}}
+       Te =  Complex{Float32}
+    else
+       Te =  Complex{Float64}
+    end
 
     if N == 1
 
@@ -622,11 +629,7 @@ function reverse_order(d::Array{Tv}; n1=0, n2=0, n3=0, n4=0, n5=0
        n1 = n1 < dims[1] ? dims[1] : n1
 
        # allocate complex array
-       if Tv <: AbstractFloat
-          r = zeros(Complex{Tv}, n1)  # input is AbstractFloat
-       else
-          r = zeros(Tv, n1)           # input is complex
-       end
+       r = zeros(Te, n1)
 
        # first dimension
        for i1 = 1 : dims[1]
[…]

### 2020-08-12 src/synthetic/get_mixture_events.jl

@@ -92,7 +92,7 @@ function get_mixture_events(; ot =0.0, dt =0.004, nt =500,
        drop     = floor(Int64, L/2)*2
 
        for ie = 1 : nevents
-           tmp = conv2(smoother, smoother, randn(nx1+drop, nx2+drop))[drop+1:end-drop, drop+1:end-drop] * sigma
+           tmp = conv(smoother, smoother, randn(nx1+drop, nx2+drop))[drop+1:end-drop, drop+1:end-drop] * sigma
            pt[ie,:,:] .= tmp[:,:]
        end
     end

### 2020-08-12 test/circulant.jl

@@ -4,8 +4,8 @@
 
       # 1D case
       N1 = 123            # length of vector
-      d  = randn(Tv, N1)  # random vector
-      v  = randn(Tv, N1)  # random vector
+      d  = randn(Tv, N1)  # randnom vector
+      v  = randn(Tv, N1)  # randnom vector
 
       C  = build_circulant_matrix(d)  # build circulant matrix
 

### 2020-08-12 test/hankel.jl

@@ -4,68 +4,74 @@
 
       # 1D case
       L1 = 12; K1 = L1-1; N1 = L1 + K1 -1 # length of vector
-      d  = randn(Tv, N1)                # random vector
+      d  = randn(Tv, N1)                # randnom vector
 
       H  = build_hankel_matrix(d)       # build hankel matrix
 
       # forward
-      v  = randn(Tv, K1)                # random vector
+      v  = randn(Tv, K1)                # randnom vector
       r1 = H * vec(v)                   # matrix times vector
-      r2 = hankel_multiplication(d, v; flag="forward")  # efficient hankel multiplication
-      r2 = vec(r2)
+      r2 = hankel_multiplication(d, v; flag="forward"); r2 = vec(r2);  # efficient hankel multiplication
+      r3 = hankel_multiplication(d, vec(v); flag="forward")
       @test norm(r1-r2)/norm(r1) < etol # test the difference
+      @test norm(r1-r3)/norm(r3) < etol # test the difference
 
       # adjoint
-      v  = randn(Tv, L1)                # random vector
-      r3 = H' * vec(v)                  # matrix times vector
-      r4 = hankel_multiplication(d, v; flag="adjoint")  # efficient hankel multiplication
-      r4 = vec(r4)
-      @test norm(r3-r4)/norm(r4) < etol # test the difference
+      v  = randn(Tv, L1)                # randnom vector
+      r1 = H' * vec(v)                  # matrix times vector
+      r2 = hankel_multiplication(d, v; flag="adjoint"); r2 = vec(r2);  # efficient hankel multiplication
+      r3 = hankel_multiplication(d, vec(v); flag="adjoint")
+      @test norm(r1-r2)/norm(r1) < etol # test the difference
+      @test norm(r1-r3)/norm(r3) < etol # test the difference
 
 
       # 2D case
       L1 = 12; K1 = L1-1; N1 = L1 + K1 -1 # length of vector
       L2 = 11; K2 = L2-1; N2 = L2 + K2 -1 # length of vector
-      d  = randn(Tv, N1, N2)            # random vector
+      d  = randn(Tv, N1, N2)            # randnom vector
 
       H  = build_hankel_matrix(d)       # build hankel matrix
 
       # forward
-      v  = randn(Tv, K1, K2)            # random vector
[…]

### 2020-08-12 test/lanczos.jl

@@ -1,0 +1,164 @@
+# ======================================================================
+#             test lanczos with partial bidiagonalization
+# ======================================================================
+using PyPlot, AcousticWave
+
+m = 100; n = 67;
+k = n;
+A = randn(m, n);
+(U, B, K) = lanbpro(A, k, r0=randn(m));
+
+m = 212; n = 100;
+path = "/Users/wenlei/Documents/AcousticHessianApprox/test/data/helio.bin"
+fid  = open(path, "r"); A = read(fid, Float64, m*n); close(fid);
+A = reshape(A, m, n);
+
+k  = n
+
+# the svd computed by lacpack
+(ls, sv_true, rs) = svd(A);
+A = ls * diagm(sv_true) * rs;
+(ls, sv, rs) = svd(A)
+
+# the relative error
+relerr = zeros(k)
+for i = 1 : k
+    relerr[i] = abs(sv[i]-sv_true[i]) / sv_true[i]
+end
+
+# computed by matrix
+r0 = rand(m) - 0.5*ones(m)
+(U, B, V) = lanbpro(A, k, r0=r0)
+
+# computed by operator
+function Amul(x, iflag; A=[])
+
+    if iflag == 1
+       y = A * x
+    elseif iflag == 2
+       y = A' * x
+    else
+       error("no third option")
+    end
+    return y
+end
+
+params = Dict(:A=>A)
+(U1, B1, V1) = lanbpro(Amul, k; m=m, n=n, params=params, r0=r0)
+
+vecnorm(U1-U)
+vecnorm(B1-B)
+vecnorm(V1-V)
+
+
+(ls1, sv1, rs1) = svd(full(B));
+relerr1 = zeros(k)
+for i = 1 : k
+    relerr1[i] = abs(sv1[i]-sv_true[i]) / sv_true[i]
+end
+plot(1:k, log10.(abs.(sv_true)), label="true")
+plot(1:k, log10.(abs.(sv     )), label="lacpack")
+plot(1:k, log10.(abs.(sv1    )), label="lanbpro")
+
+plot(1:k, log10.(relerr), label="lacpack relerr")
+plot(1:k, log10.(relerr1), label="lanbpro relerr")
+legend();
+
+# check the profile of loss orthogonality
+delta = sqrt(eps()/k)
+eta   = (eps()^(3/4)) / sqrt(k)
+
+# plot for vector V
+j = 19
[…]

### 2020-08-12 test/runtests.jl

@@ -1,4 +1,4 @@
-using Test, FFTW, LinearAlgebra
+using SeisDenoise, Test, FFTW, LinearAlgebra
 
 include("circulant.jl")
 include("toeplitz.jl")

### 2020-08-12 test/toeplitz.jl

@@ -1,4 +1,4 @@
-@testset "multi-level Hankel matrix or its adjoint times a vector" begin
+@testset "multi-level teoplitz matrix or its adjoint times a vector" begin
 
   for (Tv, etol) in [(Float32,1.0e-5), (Float64,1.0e-14), (Complex{Float32},1.0e-5), (Complex{Float64},1.0e-14)]
 
@@ -86,550 +86,5 @@
       r2    = H * vec(v_hat)
       @test norm(r1-r2)/norm(r1) < etol # test the difference
 
-  end
-end
-
-
-"""
-   slow method to count the repeated copy time when building Hankel matrix from a multi-dimensional array
-"""
-function count_copy_times_slow(dims::Union{Ti,Vector{Ti}}) where {Ti<:Int64}
-
-    # determine the dimensions of an array
-    N = length(dims)
-
-    # level 1
-    if N == 1
-
-       # allocate memory to record the number
-       count_num = zeros(Int64, dims[1])
-
-       # size of hankel matrix
-       L1 = floor(Int64,dims[1]/2)+1; K1 = dims[1]-L1+1
-
-
-       # build hankel
-       for j1 = 1 : K1
-           for i1 = 1 : L1
-               n1 = i1+j1-1
-
-               count_num[n1] += 1
-           end
-       end
-
-
-    # level 2
-    elseif N == 2
-
-       # allocate memory to record the number
-       count_num = zeros(Int64, dims[1], dims[2])
-
-       # size of hankel matrix
-       L1 = floor(Int64,dims[1]/2)+1; K1 = dims[1]-L1+1
-       L2 = floor(Int64,dims[2]/2)+1; K2 = dims[2]-L2+1
-
-       # build Hankel
-       for j2 = 1 : K2
-           c2 = (j2-1)*K1
-           for i2 = 1 : L2
-               r2 = (i2-1)*L1
-               n2 = i2 + j2 - 1
-
-               # first layer
-               for j1 = 1 : K1
-                   c1 = c2 + j1
-                   for i1 = 1 : L1
-                       r1 = r2 + i1
-                       n1 = i1 + j1 - 1
-
-                       count_num[n1,n2] += 1
-                   end
-               end #first
-           end
-       end #second
-
-
-    # level 3
-    elseif N == 3
-
-       # allocate memory to record the number
[…]
